---
title: "Residual over-dispersion and over-dispersion estimates"
author: "Nils Eling"
date: "13/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script recreates Figure 1 and Figure S1.

# Figure 1

# Compare over-dispersion and residual over-dispersion parameters

```{r}
# Read in libraries
library(BASiCS)
library(cowplot)
library(scran)
library(reshape2)

# Load MCMC chains from all datasets
MCMCs <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Testing/Tdist/All_datasets.rds")
# Select the Dictyostelium dataset
cur_MCMC <- MCMCs$MCMC_Dict_12_1.2_5_reg.rds
rm(MCMCs)

# Show the uncorrected over-dispersion values plotted against mean expression
uncorrected.fit <- BASiCS_showFit(cur_MCMC) + ylim(c(-5,12)) + xlim(c(-1, 17))

# Show the corrected residual over-dispersion values plotted against mean expression
corrected.fit <- ggplot(data.frame(mu = colMedians(cur_MCMC@parameters$mu),
                             epsilon = colMedians(cur_MCMC@parameters$epsilon))) + 
  geom_hex(aes(log(mu), epsilon), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE) +
  xlab("log(mu)") + ylab("epsilon") + theme_minimal(base_size = 15) + xlim(c(-1, 17)) +
  geom_abline(slope = 0, intercept = 0, colour="dark red")
```

# DM and epsilon values

```{r}
# Create Data object for normalization
# Read in ERCCs 
ERCC.conc <- read.table("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Data/ERCC_conc.txt", header=TRUE, sep = "\t", fill = TRUE)
ERCC.num <- matrix(data=NA, nrow=nrow(ERCC.conc), ncol=1)
ERCC.num[,1] <- (ERCC.conc[,4]*(10^(-18)))*(6.0221417*(10^23))

# Dictyostelium data
input.dict <- read.table("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Data/Test_Data/Dictyostelium.txt", sep = "\t")
# Select first time point
input.dict <- input.dict[,grepl("X0h", colnames(input.dict))]
chips <- sapply(colnames(input.dict), function(n){unlist(strsplit(n, "_"))[1]})

ERCC.num.final <- ERCC.num/1000
rownames(ERCC.num) <- rownames(ERCC.num.final) <- ERCC.conc[,2]

SpikeInput <- ERCC.num.final[rownames(input.dict)[grepl("ERCC", rownames(input.dict))],1]
SpikeInput.1 <- data.frame("Name" = names(SpikeInput),
                           "Molecules" = SpikeInput, stringsAsFactors = FALSE)

Data.dict <- newBASiCS_Data(Counts = input.dict, 
                            Tech = grepl("ERCC", rownames(input.dict)), 
                            SpikeInfo = SpikeInput.1, BatchInfo = chips)

# Normalization
Norm_counts <- BASiCS_DenoisedCounts(Data = Data.dict, Chain = cur_MCMC)

# Compute CV2
CV2 <- apply(Norm_counts[!grepl("ERCC", rownames(Norm_counts)),], 1, function(n){
  cur_var <- var(n)
  cur_mean <- mean(n)
  cur_var/(cur_mean^2)
})

# Compute DM
DM <- DM(mean = rowMeans(Norm_counts[!grepl("ERCC", rownames(Norm_counts)),]), cv2 = CV2)

# Plot DM - epsilons vs mean expression
p.Eps <- ggplot(data.frame(Mean = log(colMedians(cur_MCMC@parameters$mu)),
                           Epsilon = colMedians(cur_MCMC@parameters$epsilon),
                           DM = DM)) + 
  geom_point(aes(DM, Epsilon)) + ylim(c(-4,4)) + xlim(c(-1.5,1.5)) 
cor.test(colMedians(cur_MCMC@parameters$epsilon),DM)
ggsave("/Users/eling01/Dropbox (Personal)/BASiCSplus/Revisions/Epsilon_DM_Reviewer1.pdf",
       p.Eps)
rm(cur_MCMC)
```

# Final figure 1

```{r}
# Create final figure
Fig1 <- plot_grid(NULL, NULL, uncorrected.fit, corrected.fit, ncol = 2, nrow = 2, 
                   align = "vh", labels = "AUTO")
# Save final figure
ggsave(filename = "/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Manuscript/Figures/Revisions/Fig1.pdf", 
       plot = Fig1, width = 8, height = 8)

```

# Figure S1

The purpose of this analysis is to confirm that the EFDR holds for simulated data.
Simulations were done under the NULL model (no changes in deltas) and an 
alternative model (changes in delta for 1000 genes). We will also present the 
FPR calculated from simulations under the Null and the TPR calculated from
simulations under the alternative.

Furthermore, we validate the variability measures with matched qPCR data.

# Genrate parameters to simulate the data

```{r}
MCMC <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Testing/Tdist/All_datasets.rds")
Zeisel_MCMC <- MCMC$MCMC_Zeisel_12_1.2_5_reg.rds
rm(MCMC)
Zeisel_data <- read.table("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Data/Test_Data/microglia_Zeisel.txt")

# Collect parameters
parameters <- lapply(Zeisel_MCMC@parameters, colMedians)

# Generate mean values for spike-in information
mu_spikes <-rowMeans(Zeisel_data[grepl("ERCC", rownames(Zeisel_data)),])

# Save these information
save.image("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/DataForSimulations.RData")
```

The Simualtions are run on the cluster and scripts are available Chain generation folder.

# Visualization of the EFDR, TPR and FPR

## Figure S1 A-C

First we look at the simulations under the Null

```{r}
# Read in data - for Simulations under the Null first
FPR.mu <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/FPR_mu.rds")
FPR.delta <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/FPR_delta.rds")
FPR.eps <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/FPR_eps.rds")

EFDR.mu.Null <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/EFDR_mu_NULL.rds")
EFDR.delta.Null <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/EFDR_delta_NULL.rds")
EFDR.eps.Null <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/EFDR_eps_NULL.rds")

df <- data.frame(Values = c(FPR.mu, EFDR.mu.Null),
                 NoCells = factor(rep(c(100, 120, 200, 40, 60, 80), each = 5), 
                                  levels = c(40, 60, 80, 100, 120, 200)),
                 Statistic = rep(c("FPR", "EFDR"), each = 30))

FPR.mu <- ggplot(df) + geom_point(aes(NoCells, Values, colour = Statistic), size = 2) + 
  theme_minimal(base_size = 12) + xlab("Number of cells") + 
  scale_color_manual(values = c("#1B9E77", "#D95F02")) + 
  geom_abline(slope = 0, intercept = 0.1) + ylim(c(0,0.2)) +
  ggtitle("FPR and EFDR for changes in mean expression, Null model")

df <- data.frame(Values = c(FPR.delta, EFDR.delta.Null),
                 NoCells = factor(rep(c(100, 120, 200, 40, 60, 80), each = 5), 
                                  levels = c(40, 60, 80, 100, 120, 200)),
                 Statistic = rep(c("FPR", "EFDR"), each = 30))

FPR.delta <- ggplot(df) + geom_point(aes(NoCells, Values, colour = Statistic), size = 2) + 
  theme_minimal(base_size = 12) + xlab("Number of cells") + 
  geom_abline(slope = 0, intercept = 0.1) + 
  ylim(c(0, 0.2)) + scale_color_manual(values = c("#1B9E77", "#D95F02")) + 
  ggtitle("FPR and EFDR for changes in over-dispersion, Null model")

df <- data.frame(Values = c(FPR.eps, EFDR.eps.Null),
                 NoCells = factor(rep(c(100, 120, 200, 40, 60, 80), each = 5), 
                                  levels = c(40, 60, 80, 100, 120, 200)),
                 Statistic = rep(c("FPR", "EFDR"), each = 30))

FPR.eps <- ggplot(df) + geom_point(aes(NoCells, Values, colour = Statistic), size = 2) + 
  theme_minimal(base_size = 12) + xlab("Number of cells") + 
  geom_abline(slope = 0, intercept = 0.1) +
  ylim(c(0, 0.2)) + scale_color_manual(values = c("#1B9E77", "#D95F02")) + 
  ggtitle("FPR and EFDR for changes in residual over-dispersion, Null model")
```

## Figure S1 D+E

Now the simulations under the alternative

```{r}
# Read in data - for Simulations under the alternative
TPR.delta <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/TPR_delta_alternative.rds")
TPR.eps <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/TPR_epsilon_alternative.rds")

EFDR.delta.alternative <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/EFDR_delta_alternative.rds")
EFDR.eps.alternative <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/EFDR_eps_alternative.rds")

df <- data.frame(Values = c(TPR.delta, EFDR.delta.alternative),
                 NoCells = factor(rep(c(100, 120, 200, 40, 60, 80), each = 5), 
                                  levels = c(40, 60, 80, 100, 120, 200)),
                 Statistic = rep(c("TPR", "EFDR"), each = 30))

TPR.delta <- ggplot(df) + geom_point(aes(NoCells, Values, colour = Statistic), size = 2) + 
  theme_minimal(base_size = 12) + xlab("Number of cells") + 
  ylim(c(0, 1)) + scale_color_manual(values = c("#1B9E77", "#E7298A")) + 
  geom_abline(slope = 0, intercept = 0.1) + 
  ggtitle("TPR and EFDR for changes in over-dispersion, alternative model")

df <- data.frame(Values = c(TPR.eps, EFDR.eps.alternative),
                 NoCells = factor(rep(c(100, 120, 200, 40, 60, 80), each = 5), 
                                  levels = c(40, 60, 80, 100, 120, 200)),
                 Statistic = rep(c("TPR", "EFDR"), each = 30))

TPR.eps <- ggplot(df) + geom_point(aes(NoCells, Values, colour = Statistic), size = 2) + 
  theme_minimal(base_size = 12) + xlab("Number of cells") + 
  ylim(c(0, 1)) + scale_color_manual(values = c("#1B9E77", "#E7298A")) + 
  geom_abline(slope = 0, intercept = 0.1) + 
  ggtitle("TPR and EFDR for changes in residual over-dispersion, alternative model")
```

## Figure S1 F

We next calculate the actual FDR for epsilon and delta depending on mean expression

```{r}
# Read in MCMCs run on data simulated under the alternative
MCMC.A <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/MCMC_A_120_1.rds")
MCMC.B <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/MCMC_B_120_1.rds")

# Read in the information on genes that changes
changes <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Simulations/Deltas_B_120_1.rds")
changes <- colnames(MCMC.A@parameters$delta)[changes/parameters$delta != 1]

# Perform the testing
cur_test <- BASiCS_TestDE(MCMC.A, MCMC.B, Plot = FALSE, PlotOffset = FALSE,
                          GroupLabel1 = "A", GroupLabel2 = "B")

# Detected changes in over-dispersion
ggplot(cur_test$TableDisp[cur_test$TableDisp$ResultDiffDisp != "NoDiff" &
                               cur_test$TableDisp$ResultDiffDisp != "ExcludedFromTesting",]) +
  geom_point(aes(log(MeanOverall), DispLog2FC, 
                 colour = cur_test$TableDisp$GeneName[cur_test$TableDisp$ResultDiffDisp != "NoDiff" &
                               cur_test$TableDisp$ResultDiffDisp != "ExcludedFromTesting"] %in% changes)) + scale_color_discrete(name = "Simulated change")

ggplot(cur_test$TableResDisp[cur_test$TableResDisp$ResultDiffResDisp != "NoDiff" &
                               cur_test$TableResDisp$ResultDiffResDisp != "ExcludedFromTesting",]) +
  geom_point(aes(log(MeanOverall), ResDispDistance, 
                 colour = cur_test$TableResDisp$GeneName[cur_test$TableResDisp$ResultDiffResDisp != "NoDiff"] %in% changes)) + scale_color_discrete(name = "Simulated change")

FDR <- function(results, names, changes){
  sum(results != "NoDiff" & results != "ExcludedFromTesting" & !(names %in% changes))/
             sum(results != "NoDiff" & results != "ExcludedFromTesting")
}

mat <- matrix(data = NA, ncol = 3, nrow = 100)
colnames(mat) <- c("Mean", "FDR_delta", "FDR_eps")
mat[,1] <- seq(min(log(cur_test$TableDisp$MeanOverall)), 
               max(log(cur_test$TableDisp$MeanOverall)), 
               length.out = 100)

for(i in 1:100){
  mat[i,2] <- FDR(cur_test$TableDisp$ResultDiffDisp[log(cur_test$TableDisp$MeanOverall) > mat[i,1]],
                  cur_test$TableDisp$GeneName[log(cur_test$TableDisp$MeanOverall) > mat[i,1]],
                  changes[changes %in% 
                            cur_test$TableDisp$GeneName[log(cur_test$TableDisp$MeanOverall) > mat[i,1]]])
  mat[i,3] <- FDR(cur_test$TableResDisp$ResultDiffResDisp[log(cur_test$TableResDisp$MeanOverall) > mat[i,1]],
                  cur_test$TableResDisp$GeneName[log(cur_test$TableResDisp$MeanOverall) > mat[i,1]],
                  changes[changes %in% 
                            cur_test$TableResDisp$GeneName[log(cur_test$TableResDisp$MeanOverall) > mat[i,1]]])
}

mat.melt <- melt(as.data.frame(mat), id.vars = "Mean")

FDR <- ggplot(as.data.frame(mat.melt[mat.melt[,1] < 4,])) +
  geom_smooth(aes(Mean, value, colour = variable)) +
  scale_color_manual(values = c("dark red", "dark blue")) + 
  xlab("Mean Expression") + ylab("FDR")

```

# Final figure 1

```{r}
# Create final figure
FigS1 <- plot_grid(FPR.mu, FPR.delta, FPR.eps, 
                   TPR.delta, TPR.eps, FDR, ncol = 3, nrow = 2, 
                   align = "vh", labels = "AUTO")
# Save final figure
ggsave(filename = "/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Manuscript/Figures/Revisions/FigS1.pdf", 
       plot = FigS1, width = 12, height = 8)

```
