---
title: "Estimation stability for model parameters."
author: "Nils Eling"
date: "12/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script recreates Figure 3 and Figure S3.

Figure 3:
CA1 pyramidal neuron cells were subsetted to 50-500 cells and the regression 
as well as non-regression model were used to estimate parameters. 
These parameters are stored in one .rds file where each column represents either
mu, delta or epsilon parameters for a given sample size and iterartion 
(see script Downsampling_regression.R and Downsampling_non-regression).
We will firstly show the actual parameters for one subsampling experiment in
form of boxplots.

We will use the abbreviation NR for non-regression BASiCS model and R for
regression BASiCS model.

Figure S3:
Estimated parameters were directly compared between the regression and non-regression
model
Parameters estimated on 939 cells is used as pseudo ground truth for comparison.
We quantify the difference between estimates and the pseudo ground truth in form
of log2 fNR changes (for $\delta_i$ and $\mu_i$) and absolute distances (for $\epsilon_i$)


# Figure 3

```{r}
# Read in libraries and data
library(cowplot)
library(BASiCS)
library(plyr)
library(reshape2)
library(statmod)
library(RColorBrewer)

# Read in the paramter estimates using the NR and R model
Data.R <- as.data.frame(readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Testing/Tdist/Downsampling_Zeisel.rds"))
Data.NR <- as.data.frame(readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Testing/Tdist/Downsampling_Zeisel_old.rds"))
```

```{r}
# Offset correction between each dataset and the largest (939 cells) dataset
Offset <- function(mu_real, mu_sim){
  OS <- sum(mu_real)/sum(mu_sim) 
  mu_sim = mu_sim * OS
  return(mu_sim)
}

# Pseudo ground truth for mu
mu_real <- Data.R$MCMC_939_1.rds_mu

Data.R[,seq(1,213, 3)] <- apply(Data.R[,seq(1,213, 3)], 2, function(n){
  Offset(mu_real, n)
})
Data.NR[,seq(1,142, 2)] <- apply(Data.NR[,seq(1,142, 2)], 2, function(n){
  Offset(mu_real, n)
})

```

Visualize parameters for few samples sizes for one sub-sampling experiment.

## Figure 3 A

```{r}
# Split genes into lowly, medium and highly expressing for R and NR model
df.mu.1rep <- cbind(Data.NR[,seq(1, 142, 20)], Data.R[,seq(1, 213, 30)])
colnames(df.mu.1rep) <- c("100_NR", "150_NR", "200_NR", "250_NR", 
                             "300_NR", "500_NR", "50_NR", "939_NR",
                             "100_R", "150_R", "200_R", "250_R", 
                             "300_R", "500_R", "50_R", "939_R")
df.mu.1rep.melt <- melt(df.mu.1rep)

# We split the genes at the 33% and 66% quantile
lower <- quantile(log(mu_real), 0.33)
higher <- quantile(log(mu_real), 0.66)

df.mu.1rep.melt$ExpressionLevel <- factor(rep(ifelse(log(mu_real) < lower, "Low", 
                                             ifelse(log(mu_real) < higher, "Medium", "High")), 
                                      ncol(df.mu.1rep)),
                                  levels = c("High", "Medium", "Low"))

df.mu.1rep.melt$NoCells <- factor(as.numeric(sapply(as.character(df.mu.1rep.melt$variable), 
                                         function(n){unlist(strsplit(n, "_"))[1]})),
                                  levels = c(50, 100, 150, 200, 250, 300, 500, 939))
df.mu.1rep.melt$Model <- factor(c(rep("Non-regression", 8*nrow(df.mu.1rep)), 
                                  rep("Regression", 8*nrow(df.mu.1rep))),
                                 levels = c("Non-regression", "Regression"))

# Plot results for few sample sizes
single.mu <- ggplot(df.mu.1rep.melt[df.mu.1rep.melt$NoCells %in% c(50, 150, 250, 500, 939),]) + 
  geom_boxplot(aes(NoCells, log(value), fill=Model)) + 
  scale_fill_manual(values = c("steelblue4", "coral3")) + ylab("log(mu)") + 
  xlab("Number of cells") + facet_grid(ExpressionLevel~., scales = "free") + 
  theme(strip.background = element_blank(), strip.text.y = element_text(face = "bold", size = 16))
```

## Figure 3 B

```{r}
# Split genes into lowly, medium and highly expressing for R and NR model
df.delta.1rep <- cbind(Data.NR[,seq(2, 142, 20)], Data.R[,seq(2, 213, 30)])
colnames(df.delta.1rep) <- c("100_NR", "150_NR", "200_NR", "250_NR", 
                             "300_NR", "500_NR", "50_NR", "939_NR",
                             "100_R", "150_R", "200_R", "250_R", 
                             "300_R", "500_R", "50_R", "939_R")
df.delta.1rep.melt <- melt(df.delta.1rep)

# We split the genes at the 33% and 66% quantile
lower <- quantile(log(mu_real), 0.33)
higher <- quantile(log(mu_real), 0.66)

df.delta.1rep.melt$ExpressionLevel <- factor(rep(ifelse(log(mu_real) < lower, "Low", 
                                             ifelse(log(mu_real) < higher, "Medium", "High")), 
                                      ncol(df.delta.1rep)),
                                  levels = c("High", "Medium", "Low"))

df.delta.1rep.melt$NoCells <- factor(as.numeric(sapply(as.character(df.delta.1rep.melt$variable), 
                                         function(n){unlist(strsplit(n, "_"))[1]})),
                                  levels = c(50, 100, 150, 200, 250, 300, 500, 939))
df.delta.1rep.melt$Model <- factor(c(rep("Non-regression", 8*nrow(df.delta.1rep)), 
                                  rep("Regression", 8*nrow(df.delta.1rep))),
                                 levels = c("Non-regression", "Regression"))

# Plot results
single.delta <- ggplot(df.delta.1rep.melt[df.delta.1rep.melt$NoCells %in% c(50, 150, 250, 500, 939),]) + 
  geom_boxplot(aes(NoCells, log(value), fill=Model)) + 
  scale_fill_manual(values = c("steelblue4", "coral3")) + ylab("log(delta)") + 
  xlab("Number of cells") + facet_grid(ExpressionLevel~., scales = "free") + 
  theme(strip.background = element_blank(), strip.text.y = element_text(face = "bold", size = 16))
```

## Figure 3 C

```{r}
# Split genes into lowly, medium and highly expressing for R and NR model
df.epsilon.1rep <- Data.R[,seq(3, 213, 30)]
colnames(df.epsilon.1rep) <- c("100_R", "150_R", "200_R", "250_R", 
                             "300_R", "500_R", "50_R", "939_R")
df.epsilon.1rep.melt <- melt(df.epsilon.1rep)

# We split the genes at the 33% and 66% quantile
lower <- quantile(log(mu_real), 0.33)
higher <- quantile(log(mu_real), 0.66)

df.epsilon.1rep.melt$ExpressionLevel <- factor(rep(ifelse(log(mu_real) < lower, "Low", 
                                             ifelse(log(mu_real) < higher, "Medium", "High")), 
                                      ncol(df.epsilon.1rep)),
                                  levels = c("High", "Medium", "Low"))

df.epsilon.1rep.melt$NoCells <- factor(as.numeric(sapply(as.character(df.epsilon.1rep.melt$variable), 
                                         function(n){unlist(strsplit(n, "_"))[1]})),
                                  levels = c(50, 100, 150, 200, 250, 300, 500, 939))
df.epsilon.1rep.melt$Model <- factor(c(rep("Regression", 8*nrow(df.epsilon.1rep))))

# Plot results
single.epsilon <- ggplot(df.epsilon.1rep.melt[df.epsilon.1rep.melt$NoCells %in% c(50, 150, 250, 500, 939),]) + 
  geom_boxplot(aes(NoCells, value, fill=Model)) + 
  scale_fill_manual(values = c("coral3")) + ylab("epsilon")  + xlab("Number of cells") + 
  facet_grid(ExpressionLevel~., scales = "free") + 
  theme(strip.background = element_blank(), strip.text.y = element_text(face = "bold", size = 16))
```

## Save Figure 3

```{r}
Fig3 <- plot_grid(single.mu, single.delta, single.epsilon, 
                  ncol = 3, nrow = 1, align = "v", labels = "AUTO")

ggsave("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Manuscript/Figures/Revisions/Fig3.pdf", plot = Fig3, width = 24, height = 10)
```

# Figure S3

## Figure S3A

Direct comparisons between parameters estimated from the NR and R model. 
Exemplified are sample sizes of 50, 250 and 939 cell.

```{r}
# Smallest population
direct.delta.small <- ggplot(data.frame(dRvsdNR = log2(Data.R$MCMC_50_1.rds_delta/Data.NR$MCMC_50_1.rds_delta),
                  Mean = log(rowMeans(matrix(c(Data.R$MCMC_50_1.rds_mu, Data.NR$MCMC_50_1.rds_mu), 
                                             nrow = nrow(Data.R), ncol = 2,
                                             byrow = FALSE))))) + ylim(c(-4, 8)) + xlim(c(-4, 9)) +
  geom_hex(aes(Mean, dRvsdNR), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE)+
  geom_abline(slope = 0, intercept = 0, colour = "dark red", lwd = 1)

# intermediate population
direct.delta.inter <- ggplot(data.frame(
                  dRvsdNR = log2(Data.R$MCMC_250_1.rds_delta/Data.NR$MCMC_250_1.rds_delta),
                  Mean = log(rowMeans(matrix(c(Data.R$MCMC_250_1.rds_mu, Data.NR$MCMC_250_1.rds_mu), 
                                             nrow = nrow(Data.R), ncol = 2,
                                             byrow = FALSE))))) + ylim(c(-4, 8)) + xlim(c(-4, 9)) +
  geom_hex(aes(Mean, dRvsdNR), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE)+
  geom_abline(slope = 0, intercept = 0, colour = "red")

# Largest population
direct.delta.pgt <- ggplot(data.frame(dRvsdNR = log2(Data.R$MCMC_939_1.rds_delta/Data.NR$MCMC_939_1.rds_delta),
                  Mean = log(rowMeans(matrix(c(Data.R$MCMC_939_1.rds_mu, Data.NR$MCMC_939_1.rds_mu), 
                                             nrow = length(Data.R$MCMC_939_1.rds_mu), ncol = 2,
                                             byrow = FALSE))))) + ylim(c(-4, 8)) + xlim(c(-4, 9)) +
 geom_hex(aes(Mean, dRvsdNR), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE)+
  geom_abline(slope = 0, intercept = 0, colour = "red")
```

## Figure S3B

```{r}
# Split genes into lowly, medium and highly expressing for R and NR model
df.mu <- matrix(data=NA, ncol = 1, nrow = 8*30*2)
colnames(df.mu) <- c("Median")
idx <- 1
counter <- 1

# We split the genes at the 33% and 66% quantile
lower <- quantile(log(mu_real), 0.33)
higher <- quantile(log(mu_real), 0.66)

# We collect the median and the central 50% quantile
for(i in seq(1,213, 3)){
  cur_mu_R <- Data.R[,i]
  cur_mu_NR <- Data.NR[,idx]
  
  df.mu[counter,1] <- median(log2(cur_mu_NR[log(mu_real) < lower] / 
                                    mu_real[log(mu_real) < lower]))
  
  df.mu[counter+1,1] <- median(log2(cur_mu_NR[log(mu_real) > lower & log(mu_real) < higher] / 
                                      mu_real[log(mu_real) > lower & log(mu_real) < higher]))
  
  df.mu[counter+2,1] <- median(log2(cur_mu_NR[log(mu_real) > higher] / 
                               mu_real[log(mu_real) > higher]))
  
  df.mu[counter+3,1] <- median(log2(cur_mu_R[log(mu_real) < lower] / 
                                    mu_real[log(mu_real) < lower]))
  
  df.mu[counter+4,1] <- median(log2(cur_mu_R[log(mu_real) > lower & log(mu_real) < higher] / 
                                      mu_real[log(mu_real) > lower & log(mu_real) < higher]))
  
  df.mu[counter+5,1] <- median(log2(cur_mu_R[log(mu_real) > higher] / 
                               mu_real[log(mu_real) > higher]))
  counter <- counter+6
  idx = idx+2
}

# Create collapsed data.frame
df.mu.df <- as.data.frame(df.mu)
df.mu.df$NoCells <- factor(rep(c(100, 150, 200, 250, 300, 500, 50, 939), each=60), 
                             levels = c(50, 100, 150, 200, 250, 300, 500, 939))
df.mu.df$Model <- factor(c(rep(c("Non-regression", "Non-regression", "Non-regression",
                                 "Regression", "Regression", "Regression"), 80)), 
                            levels = c("Non-regression", "Regression"))
df.mu.df$ExpressionLevel <- factor(rep(c("Low", "Medium", "High"), 160), 
                            levels = c("High", "Medium", "Low"))

df.mu.df <- df.mu.df[which(!is.na(df.mu.df[,1])),]

boxplot.mu <- ggplot(df.mu.df) + 
  geom_point(aes(NoCells, Median, colour=Model), size=1, position=position_dodge(width = 0.5)) +
  scale_color_manual(values = c("steelblue4", "coral3")) + ylab("log2(mu_i/mu_i,pgt)") + 
  xlab("Number of cells") + facet_grid(ExpressionLevel~.) + geom_hline(yintercept = 0, colour = "dark red") +
  ylim(c(-1,1)) +
  theme(strip.background = element_blank(), strip.text.y = element_text(face = "bold", size = 16))
```

```{r}
# Split genes into lowly, medium and highly expressing for R and NR model
df.delta <- matrix(data=NA, ncol = 1, nrow = 8*30*2)
colnames(df.delta) <- c("Median")
idx <- 2
counter <- 1

# Pseudo ground truth for delta
delta_real <- Data.R$MCMC_939_1.rds_delta

# We split the genes at the 33% and 66% quantile
lower <- quantile(log(mu_real), 0.33)
higher <- quantile(log(mu_real), 0.66)

# We collect the median and the central 50% quantile
for(i in seq(2,213, 3)){
  cur_delta_R <- Data.R[,i]
  cur_delta_NR <- Data.NR[,idx]
  
  df.delta[counter,1] <- median(log2(cur_delta_NR[log(mu_real) < lower] / 
                                    delta_real[log(mu_real) < lower]))
  
  df.delta[counter+1,1] <- median(log2(cur_delta_NR[log(mu_real) > lower & log(mu_real) < higher] / 
                                      delta_real[log(mu_real) > lower & log(mu_real) < higher]))
  
  df.delta[counter+2,1] <- median(log2(cur_delta_NR[log(mu_real) > higher] / 
                               delta_real[log(mu_real) > higher]))
  
  df.delta[counter+3,1] <- median(log2(cur_delta_R[log(mu_real) < lower] / 
                                    delta_real[log(mu_real) < lower]))
  
  df.delta[counter+4,1] <- median(log2(cur_delta_R[log(mu_real) > lower & log(mu_real) < higher] / 
                                      delta_real[log(mu_real) > lower & log(mu_real) < higher]))
  
  df.delta[counter+5,1] <- median(log2(cur_delta_R[log(mu_real) > higher] / 
                               delta_real[log(mu_real) > higher]))
  counter <- counter+6
  idx = idx+2
}

# Create collapsed data.frame
df.delta.df <- as.data.frame(df.delta)
df.delta.df$NoCells <- factor(rep(c(100, 150, 200, 250, 300, 500, 50, 939), each=60), 
                             levels = c(50, 100, 150, 200, 250, 300, 500, 939))
df.delta.df$Model <- factor(c(rep(c("Non-regression", "Non-regression", "Non-regression",
                                 "Regression", "Regression", "Regression"), 80)), 
                            levels = c("Non-regression", "Regression"))
df.delta.df$ExpressionLevel <- factor(rep(c("Low", "Medium", "High"), 160), 
                            levels = c("High", "Medium", "Low"))

df.delta.df <- df.delta.df[which(!is.na(df.delta.df[,1])),]

boxplot.delta <- ggplot(df.delta.df) + 
  geom_point(aes(NoCells, Median, colour=Model), size=1, position=position_dodge(width = 0.5)) +
  scale_color_manual(values = c("steelblue4", "coral3")) + ylab("log2(delta_i/delta_i,pgt)") + 
  xlab("Number of cells") + facet_grid(ExpressionLevel~.) + geom_hline(yintercept = 0, colour = "dark red") +
  ylim(c(-1,1)) +
  theme(strip.background = element_blank(), strip.text.y = element_text(face = "bold", size = 16))
```

```{r}
mu_real <- Data.R$MCMC_939_1.rds_mu
epsilon_real <- Data.R$MCMC_939_1.rds_epsilon
df.epsilon <- matrix(data=NA, ncol = 1, nrow = 8*30)
colnames(df.epsilon) <- c("Median")
counter <- 1

# We collect the median and the central 50% quantile
for(i in seq(3,213, 3)){
  cur_epsilon <- Data.R[,i]
  
  df.epsilon[counter,1] <- median(cur_epsilon[log(mu_real) < lower] - 
                                    epsilon_real[log(mu_real) < lower], na.rm = TRUE)
  
  df.epsilon[counter+1,1] <- median(cur_epsilon[log(mu_real) > lower & log(mu_real) < higher] -
                                     epsilon_real[log(mu_real) > lower & log(mu_real) < higher], na.rm = TRUE)
  
  df.epsilon[counter+2,1] <- median(cur_epsilon[log(mu_real) > higher] -
                                      epsilon_real[log(mu_real) > higher], na.rm = TRUE)
  
  counter <- counter+3
}

# Create collapsed data.frame
df.epsilon.df <- as.data.frame(df.epsilon)
df.epsilon.df$NoCells <- factor(rep(c(100, 150, 200, 250, 300, 500, 50, 939), each=30), 
                             levels = c(50, 100, 150, 200, 250, 300, 500, 939))
df.epsilon.df$ExpressionLevel <- factor(rep(c("Low", "Medium", "High"), 80), 
                            levels = c("High", "Medium", "Low"))

df.epsilon.df <- df.epsilon.df[which(!is.na(df.epsilon.df[,1])),]

boxplot.epsilon <- ggplot(df.epsilon.df) + 
  geom_point(aes(NoCells, Median), colour="coral3", size=1, position=position_dodge(width = 0.5)) +
  ylab("epsilon_i - epsilon_i,pgt)") + ylim(c(-0.2, 0.2)) +
  xlab("Number of cells") + facet_grid(ExpressionLevel~.) + geom_hline(yintercept = 0, colour = "dark red") +
  theme(strip.background = element_blank(), strip.text.y = element_text(face = "bold", size = 16))
```

## Figure S3C-G

Compare noise measures between scRNAseq and smFISH data.

```{r}
# Read in MCMC chains for Grun data
# These were run independenyly for following conditions: SC_2i, SC_serum, RNA_2i and RNA_serum
MCMCs <- readRDS("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/qPCRvalidation/data/MCMC_ps_all.rds")

# smFISH data
files <- list.files("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/qPCRvalidation/data/smFISH_counts/", full.names = TRUE)
file.names <- list.files("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/qPCRvalidation/data/smFISH_counts/", full.names = FALSE)
smFISH.counts <- lapply(files, function(n){
  cur_file <- read.table(n, header = TRUE)
})

# Collect mean expression and CV2 measures of smFISH data in one dataframe
df.FISH <- data.frame(genenames = sapply(sapply(file.names, function(n){unlist(strsplit(n, "_"))[4]}),
                                    function(x){unlist(strsplit(x, "\\."))[1]}),
                 condition = sapply(file.names, function(n){unlist(strsplit(n, "_"))[3]}),
                 mean = unlist(lapply(smFISH.counts, function(n){mean(n[,1])})),
                 CV2 = unlist(lapply(smFISH.counts, function(n){var(n[,1])/(mean(n[,1]))^2})))

# We remove Notch1 due to strong disagreement between scRNAseq and smFISH in the serum condition.
df.FISH <- df.FISH[df.FISH$genenames != "Notch1",]

# Plot CV2 vs mean expression
ggplot(df.FISH) + geom_point(aes(log(mean), CV2, colour = condition))

# Fit reciprocal gamma model
fit.gam <- glmgam.fit(y = df.FISH$CV2, X = matrix(1/log(df.FISH$mean), ncol = 1, nrow = 16))
res <- df.FISH$CV2 - fit.gam$fitted.values

# Perform differential testing between serum and 2i condition using scRNAseq data
Test.SC <- BASiCS_TestDE(Chain1 = MCMCs$SC_2i, Chain2 = MCMCs$SC_serum,
                         GroupLabel1 = "2i", GroupLabel2 = "Serum",
                         Plot = FALSE, PlotOffset = FALSE)

# Compare model parameters vs mean, CV2 and residual CV2
genenames <- as.character(sapply(sapply(file.names, function(n){unlist(strsplit(n, "_"))[4]}),
                                    function(x){unlist(strsplit(x, "\\."))[1]}))
genenames <- genenames[genenames != "Notch1"]
df.comp <- data.frame(genenames = genenames,
                 means = rowMeans(matrix(data = df.FISH$mean, ncol = 2, nrow = 8, byrow = FALSE)),
                 Condition = df.FISH$condition,
                 FISHmean = df.FISH$mean,
                 FISHCV2 = df.FISH$CV2,
                 residual = res,
                 mu = c(Test.SC$TableMean$Mean1[match(genenames[1:8],Test.SC$TableMean$GeneName)], 
                        Test.SC$TableMean$Mean2[match(genenames[1:8],Test.SC$TableMean$GeneName)]),
                 delta = c(Test.SC$TableDisp$Disp2[match(genenames[1:8],Test.SC$TableMean$GeneName)], 
                        Test.SC$TableDisp$Disp2[match(genenames[1:8],Test.SC$TableMean$GeneName)]),
                 eps = c(Test.SC$TableResDisp$ResDisp1[match(genenames[1:8],Test.SC$TableMean$GeneName)], 
                        Test.SC$TableResDisp$ResDisp2[match(genenames[1:8],Test.SC$TableMean$GeneName)]))

smFISH.mean <- ggplot(df.comp) + geom_point(aes(log(FISHmean), log(mu), colour = Condition), size = 3) +
  geom_abline(slope = 1, intercept = 0, colour = "dark red", lwd = 1) + ylim(c(0, 7)) + xlim(c(0,7)) +
  scale_colour_brewer(palette = "Set1") 
cor.test(log(df.comp$FISHmean), log(df.comp$mu))

smFISH.CV2 <- ggplot(df.comp) + geom_point(aes(log(FISHCV2), log(delta), colour = Condition), size = 3) +
  geom_abline(slope = 1, intercept = 0, colour = "dark red", lwd = 1) + ylim(c(-3, 1)) + xlim(c(-3,1)) +
  scale_colour_brewer(palette = "Set1") 
cor.test(log(df.comp$FISHCV2), log(df.comp$delta))

smFISH.rCV2 <- ggplot(df.comp) + geom_point(aes(res, eps, colour = Condition), size = 3) + theme_minimal() + 
  ylim(c(-2,2)) + xlim(c(-0.5, 0.5)) + geom_abline(slope = 0, intercept = 0, colour = "dark blue", lwd = 1) +
  scale_colour_brewer(palette = "Set1") + geom_vline(xintercept = 0, colour = "dark blue", lwd = 1)
cor.test(df.comp$res, df.comp$eps)
```

We now compare the detection of differential variability.

```{r}
df.changes <- data.frame(genenames = df.FISH$genenames[df.FISH$condition == "J12i"],
                         mean = rowMeans(matrix(data = df.FISH$mean, ncol = 2, nrow = 8, byrow = FALSE)),
                         changes_CV2 = log2(df.FISH$CV2[df.FISH$condition == "J12i"]/df.FISH$CV2[df.FISH$condition == "J1S"]),
                         DV_results = Test.SC$TableDisp$ResultDiffDisp[
                           match(df.FISH$genenames[df.FISH$condition == "J12i"],Test.SC$TableDisp$GeneName)],
                         rDV_results = Test.SC$TableResDisp$ResultDiffResDisp[
                           match(df.FISH$genenames[df.FISH$condition == "J12i"],Test.SC$TableResDisp$GeneName)]) 

# Plot Changes in variability vs mean expression
FISH.DV.delta <- ggplot(df.changes, aes(log(mean), changes_CV2, label = genenames)) + 
  geom_point() + geom_label(aes(fill = factor(DV_results)), colour = "black", fontface = "bold") +
  scale_fill_manual(values = c("white", "red")) + ylim(c(-1.5, 1.5)) + ggtitle("DV testing using over-dispersion")

FISH.DV.eps <- ggplot(df.changes, aes(log(mean), changes_CV2, label = genenames)) + 
  geom_point() + geom_label(aes(fill = factor(rDV_results)), colour = "black", fontface = "bold") +
  scale_fill_manual(values = c("white", "red")) + ylim(c(-1.5, 1.5)) + 
  ggtitle("DV testing using residual over-dispersion")

```

## Save Figure S3

```{r}
FigS3 <- plot_grid(direct.delta.small, direct.delta.inter, direct.delta.pgt,
                   boxplot.mu, boxplot.delta, boxplot.epsilon,
                   smFISH.mean, smFISH.CV2, smFISH.rCV2, 
                   FISH.DV.delta, FISH.DV.eps, NULL, 
                   ncol = 3, nrow = 4, align = "hv", labels = "AUTO")

ggsave("/Users/eling01/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Manuscript/Figures/Revisions/FigS3.pdf", plot = FigS3, width = 20, height = 20)
```
