---
title: "CD4 T cell activation"
author: "nils Eling"
date: "18/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script reproduces Figure 5 and Figure S5.

# Figure 5

This analysis uses MCMC runs on naive and activated CD4$^+$ T cells from young B6.
Data from Martinez et al. was used to find genes that change in variability 
during early immune activation.

Here we show the advantages of the BASiCS extension for 
differential variability testing.

```{r}
# Load libraries and extra data
library(BASiCS)
library(cowplot)
library(reshape2)
library(pheatmap)
library(ggplot2)
library(viridis)
library(VennDiagram)

# Read in mouse genenames  
mouse.genes <- read.table("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Data/Genenames.txt", sep = '\t', header = TRUE, 
                          stringsAsFactors = FALSE)
rownames(mouse.genes) <- mouse.genes[,1]

# Load chains run on naive and activated CD4 T cells
MCMCs <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/All_CD4.rds")

# Exclude lowly expressed genes 
genes <- (colMedians(MCMCs$YoungNaive@parameters$mu) > 1 &
            colMedians(MCMCs$YoungActive@parameters$mu) > 1)

# Perform differential testing between naive and activated CD4+ T cells

Test.NaiveActive <- BASiCS_TestDE(Chain1 = MCMCs$YoungActive, Chain2 = MCMCs$YoungNaive, 
                                  EpsilonM = 1, 
                                  OffSet = TRUE, GenesSelect = genes, Plot = FALSE, 
                                  PlotOffset = FALSE)

# Exlude the lowly expressed genes
Test.NaiveActive$TableResDisp <- Test.NaiveActive$TableResDisp[
  Test.NaiveActive$TableResDisp$ResultDiffResDisp != "ExcludedByUser",]

Test.NaiveActive$TableMean <- Test.NaiveActive$TableMean[
  match(Test.NaiveActive$TableResDisp$GeneName, Test.NaiveActive$TableMean$GeneName),]
```

## Figure 5 A

We can now separate the data into four categories: 
* up-regulated with decrease in variability
* up-regulated with increase in variability
* down-regulated with decrease in variability
* down-regulated with increase in variability

```{r}
# Form a data.frame to collect genes in each group
df <- data.frame(row.names = Test.NaiveActive$TableMean$GeneName,
                 DistEpsilon = Test.NaiveActive$TableResDisp$ResDispDistance,
                 Log2FCmu = Test.NaiveActive$TableMean$MeanLog2FC,
                 Regulation = paste(Test.NaiveActive$TableMean$ResultDiffMean,
                               Test.NaiveActive$TableResDisp$ResultDiffResDisp, sep = "_"))
df$Regulation <- ifelse(df$Regulation == "Group1+_Group1+", "Up-regulated, higher variability", 
                   ifelse(df$Regulation == "Group1+_Group2+", "Up-regulated, lower variability",
                          ifelse(df$Regulation == "Group2+_Group2+", "Down-regulated, lower variability",
                                 ifelse(df$Regulation == "Group2+_Group1+", "Down-regulated, higher variability", "Other"))))

# Plot the figure
fourGroups <- ggplot(df, aes(Log2FCmu, DistEpsilon)) + geom_point(aes(colour=Regulation)) + 
  xlab("Log2FC mu") + ylab("Distance epsilon") + 
  scale_colour_manual(values = c("violetred4", "violetred2", "grey90", "seagreen2", "seagreen4"))
```

But we can also look at single genes.

We select a few genes to represent the regulatory mechanisms behind CD4+ T cell activation

```{r}
#### Read in data
# We need to normalize data before plotting

input <- read.table("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Data/Test_Data/CD4_NaiveActiveYoungB6.txt", sep = "\t")

#### Read in Spike-ins

ERCC.conc <- read.table("../Data/ERCC_conc.txt", header=TRUE, sep = "\t", fill = TRUE)

ERCC.num <- matrix(data=NA, nrow=nrow(ERCC.conc), ncol=1)
ERCC.num[,1] <- (ERCC.conc[,4]*(10^(-18)))*(6.0221417*(10^23))
ERCC.num.final <- ERCC.num/50000
rownames(ERCC.num) <- rownames(ERCC.num.final) <- ERCC.conc[,2]

SpikeInput <- ERCC.num.final[rownames(input)[grepl("ERCC", rownames(input))],1]
SpikeInput.1 <- data.frame("Name" = names(SpikeInput),
                           "Molecules" = SpikeInput,
                           stringsAsFactors = FALSE)

#### Create Data objects for each condition

# Young naive B6
cur_in <- input[,grepl("Unstimulated", colnames(input))]
cur_chips <- sapply(colnames(cur_in), function(n){unlist(strsplit(n, "_"))[1]})

Data.YoungNaiveB6 <- newBASiCS_Data(Counts = cur_in, 
                                    Tech = grepl("ERCC", rownames(input)), 
                                    SpikeInfo = SpikeInput.1, 
                                    BatchInfo = cur_chips)

# Young active B6
cur_in <- input[,grepl("Active", colnames(input))]
cur_chips <- sapply(colnames(cur_in), function(n){unlist(strsplit(n, "_"))[1]})

Data.YoungActiveB6 <- newBASiCS_Data(Counts = cur_in, 
                                    Tech = grepl("ERCC", rownames(input)), 
                                    SpikeInfo = SpikeInput.1, 
                                    BatchInfo = cur_chips)

# Offset correction
OffSet <- function(MCMC1, MCMC2){
  OS <- median(rowSums(MCMC1@parameters$mu)/rowSums(MCMC2@parameters$mu))
  OS
}

OS <- OffSet(MCMCs$YoungNaive, MCMCs$YoungActive)

MCMCs$YoungNaive@parameters$mu <- MCMCs$YoungNaive@parameters$mu / OS
MCMCs$YoungNaive@parameters$phi <- MCMCs$YoungNaive@parameters$phi * OS

Counts.naive <- BASiCS_DenoisedCounts(Data.YoungNaiveB6, MCMCs$YoungNaive)
Counts.active <- BASiCS_DenoisedCounts(Data.YoungActiveB6, MCMCs$YoungActive)
```

## Figure 5 B

Now we can visualize single genes.
First we represent genes that decrease in variability and become up0regulated 
during immune activation.

```{r}
rownames(Counts.active) <- mouse.genes[rownames(Counts.active),2]
rownames(Counts.naive) <- mouse.genes[rownames(Counts.naive),2]

genes <- c("Polr2l", "Pes1", "Mapkapk3", "Map3k8", "Ksr1")

df <- data.frame(value = c(as.vector(t(Counts.active[genes,])), as.vector(t(Counts.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(Counts.active)),
                                  rep(genes, each = ncol(Counts.naive))), levels = rev(genes)),
                 Condition = factor(c(rep("Active", length(as.vector(t(Counts.active[genes,])))),
                                      rep("Naive", length(as.vector(t(Counts.naive[genes,]))))), 
                                    levels = c("Active", "Naive"))
                 )

final.lessVar <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition)) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + 
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))

genes <- c("Fasl", "Smad3",
  "Pou2f2", "Il2", "Cd274")

df <- data.frame(value = c(as.vector(t(Counts.active[genes,])), as.vector(t(Counts.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(Counts.active)),
                                  rep(genes, each = ncol(Counts.naive))), levels = rev(genes)),
                 Condition = factor(c(rep("Active", length(as.vector(t(Counts.active[genes,])))),
                                      rep("Naive", length(as.vector(t(Counts.naive[genes,]))))), 
                                    levels = c("Active", "Naive"))
                 )

final.moreVar <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition)) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + 
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
```

## Figure 5 C

Now look at genes that increase in variability and become up-regulated during 
immune activation.

```{r}
genes <- c("Fasl", "Smad3",
  "Pou2f2", "Il2", "Cd274")

df <- data.frame(value = c(as.vector(t(Counts.active[genes,])), as.vector(t(Counts.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(Counts.active)),
                                  rep(genes, each = ncol(Counts.naive))), levels = rev(genes)),
                 Condition = factor(c(rep("Active", length(as.vector(t(Counts.active[genes,])))),
                                      rep("Naive", length(as.vector(t(Counts.naive[genes,]))))), 
                                    levels = c("Active", "Naive"))
                 )

final.moreVar <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition)) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + 
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
```

Save final figure.

```{r}
# Build the whole figure
Fig5 <- plot_grid(fourGroups, 
                   plot_grid(final.lessVar,
                             final.moreVar, 
                             ncol=2, nrow=1, 
                             align = "h", labels=c("B", "C")), 
                   ncol = 1, nrow = 2, labels = c("A", NA))

ggsave("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Manuscript/Figures/Fig3.pdf", plot = Fig5, width = 7, height = 10)

```

# Figure S5

## Figure S5 A+B

Comparison of DV genes detected by regression and non-regression model

```{r}
# Load extra data

# Load chains of active and naive T cells generated by regression and 
# non-regression case
MCMC.naive.new <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Old_new_comparison_naiveactive/MCMC_CD4_naive_12_1.2_5_reg.rds")
MCMC.naive.old <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Old_new_comparison_naiveactive/MCMC_CD4_naive_old.rds")
MCMC.active.new <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Old_new_comparison_naiveactive/MCMC_CD4_active_12_1.2_5_reg.rds")
MCMC.active.old <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Old_new_comparison_naiveactive/MCMC_CD4_active_old.rds")

# Perform testing between naive and active populations for regression and non-regression case
# First exclude genes that don't show at least mu=1 in one of the conditions
genes <- (colMedians(MCMC.naive.new@parameters$mu) > 50 |
            colMedians(MCMC.active.new@parameters$mu) > 50) &
  (colMedians(MCMC.naive.old@parameters$mu) > 50 |
            colMedians(MCMC.active.old@parameters$mu) > 50)

# Perform testing on log2 FC = 2 as in Martinez-Jimenez et al, 2017 to detect changes in mean expression
Test_Regression.DE <- BASiCS_TestDE(MCMC.naive.new, MCMC.active.new, EpsilonM = 2,
                                 Plot = FALSE, PlotOffset = FALSE,
                                 GroupLabel1 = "Naive_Regression", 
                                 GroupLabel2 = "Active_Regression",
                                 GenesSelect = genes)

Test_NonRegression.DE <- BASiCS_TestDE(MCMC.naive.old, MCMC.active.old, EpsilonM = 2,
                                 Plot = FALSE, PlotOffset = FALSE,
                                 GroupLabel1 = "Naive_NonRegression", 
                                 GroupLabel2 = "Active_NonRegression",
                                 GenesSelect = genes)

# Perform testing on log2 FC = 0 as in Martinez-Jimenez et al, 2017 to detect changes in varaibility
Test_Regression.DV <- BASiCS_TestDE(MCMC.naive.new, MCMC.active.new, EpsilonM = 0,
                                 Plot = FALSE, PlotOffset = FALSE,
                                 GroupLabel1 = "Naive_Regression", 
                                 GroupLabel2 = "Active_Regression",
                                 GenesSelect = genes)

Test_NonRegression.DV <- BASiCS_TestDE(MCMC.naive.old, MCMC.active.old, EpsilonM = 0,
                                 Plot = FALSE, PlotOffset = FALSE,
                                 GroupLabel1 = "Naive_NonRegression", 
                                 GroupLabel2 = "Active_NonRegression",
                                 GenesSelect = genes)

# Figure S5A
# First we compare the differential mean expression test between the two models
# Genes that are detected as differentially expressed
NumberDE.Regression <- sum(Test_Regression.DE$TableMean$ResultDiffMean != "NoDiff" &
                             Test_Regression.DE$TableMean$ResultDiffMean != "ExcludedByUser")
NumberDE.NonRegression <- sum(Test_NonRegression.DE$TableMean$ResultDiffMean != "NoDiff" &
                                Test_NonRegression.DE$TableMean$ResultDiffMean != "ExcludedByUser")
Overlap <- length(intersect(Test_Regression.DE$TableMean$GeneName[
                                  Test_Regression.DE$TableMean$ResultDiffMean != "NoDiff" &
                                    Test_Regression.DE$TableMean$ResultDiffMean != "ExcludedByUser"],
                            Test_NonRegression.DE$TableMean$GeneName[
                                  Test_NonRegression.DE$TableMean$ResultDiffMean != "NoDiff" &
                                    Test_NonRegression.DE$TableMean$ResultDiffMean != "ExcludedByUser"]
                         ))
draw.pairwise.venn(area1 = NumberDE.NonRegression, area2 = NumberDE.Regression, cross.area = Overlap, 
                   category = c("Differential mean: Non Regression case", 
                                "Differential mean: Regression case"), euler.d = TRUE, scaled = TRUE,
                   fill = c("steelblue4", "coral3"))

# Figure S5B
# Genes that are not detected as differentially expressed
NumberDE.Regression <- sum(Test_Regression.DV$TableMean$ResultDiffMean == "NoDiff")
NumberDE.NonRegression <- sum(Test_NonRegression.DV$TableMean$ResultDiffMean == "NoDiff")
Overlap <- length(intersect(Test_Regression.DV$TableMean$GeneName[
                                  Test_Regression.DV$TableMean$ResultDiffMean == "NoDiff"],
                            Test_NonRegression.DV$TableMean$GeneName[
                                  Test_NonRegression.DV$TableMean$ResultDiffMean == "NoDiff"]
                         ))
draw.pairwise.venn(area1 = NumberDE.NonRegression, area2 = NumberDE.Regression, cross.area = Overlap, 
                   category = c("No differential mean: Non Regression case", 
                                "No differential mean: Regression case"), euler.d = TRUE, scaled = TRUE,
                   fill = c("steelblue4", "coral3"))

# Second, we compare the DV genes between regression and non-regression case using the deltas
# Select the genes that are not differentially expressed in both comparisons
genes <- intersect(Test_Regression.DV$TableMean$GeneName[
                                  Test_Regression.DV$TableMean$ResultDiffMean == "NoDiff"],
                            Test_NonRegression.DV$TableMean$GeneName[
                                  Test_NonRegression.DV$TableMean$ResultDiffMean == "NoDiff"]
                         )
NumberDV.Regression <- sum(Test_Regression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                         Test_Regression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_Regression.DV$TableDisp$GeneName %in% genes)
NumberDV.NonRegression <- sum(Test_NonRegression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                         Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_NonRegression.DV$TableDisp$GeneName %in% genes)
Overlap <- length(intersect(Test_Regression.DV$TableDisp$GeneName[
                                  Test_Regression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                                  Test_Regression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_Regression.DV$TableDisp$GeneName %in% genes
                                  ],
                            Test_NonRegression.DV$TableDisp$GeneName[
                                  Test_NonRegression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                                  Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_NonRegression.DV$TableDisp$GeneName %in% genes
                                  ]
                         ))

# Jaccard index
Overlap/(NumberDV.Regression + NumberDV.NonRegression - Overlap)

draw.pairwise.venn(area1 = NumberDV.NonRegression, area2 = NumberDV.Regression, cross.area = Overlap, 
                   category = c("Differential dispersion :Non Regression case", 
                                "Differential dispersion: Regression case"), euler.d = TRUE, scaled = TRUE,
                   fill = c("steelblue4", "coral3"))

# Number of genes that can be tested for: 1. differential over-dispersion and
#                                         2. differential residual over-dispersion
sum(Test_Regression.DV$TableResDisp$ResultDiffResDisp != "ExcludedFromTesting" & 
                           Test_Regression.DV$TableResDisp$ResultDiffResDisp != "ExcludedByUser")
sum(Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedByUser")
```

# Compare analysis across batches

```{r}
MCMC.batch1.naive <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Single_batches/MCMC_naive_SS51.rds")
MCMC.batch1.active <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Single_batches/MCMC_active_SS51.rds")
MCMC.batch2.naive <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Single_batches/MCMC_naive_SS52.rds")
MCMC.batch2.active <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Single_batches/MCMC_active_SS52.rds")

genes <- (colMedians(MCMC.batch1.naive@parameters$mu) > 1 &
            colMedians(MCMC.batch1.active@parameters$mu) > 1 &
            colMedians(MCMC.batch2.naive@parameters$mu) > 1 &
            colMedians(MCMC.batch2.active@parameters$mu) > 1)

Test1 <- BASiCS_TestDE(MCMC.batch1.naive, MCMC.batch1.active,
                                 Plot = FALSE, PlotOffset = FALSE,
                                 GroupLabel1 = "Naive1", EpsilonM = 1, GenesSelect = genes,
                                 GroupLabel2 = "Active1")

Test2 <- BASiCS_TestDE(MCMC.batch2.naive, MCMC.batch2.active,
                                 Plot = FALSE, PlotOffset = FALSE,
                                 GroupLabel1 = "Naive2", EpsilonM = 1, GenesSelect = genes,
                                 GroupLabel2 = "Active2")

# Visualize correlation of residuals between batches
# Naive cells
ggplot(data.frame(mean = log(rowMeans(matrix(c(colMedians(MCMC.batch1.naive@parameters$mu[]),
                                               colMedians(MCMC.batch2.naive@parameters$mu)),
                                             ncol = 2, 
                                             nrow = length(colMedians(MCMC.batch1.naive@parameters$mu)),
                                             byrow = FALSE))),
  distance = colMedians(MCMC.batch1.naive@parameters$epsilon) -
    colMedians(MCMC.batch2.naive@parameters$epsilon))) +
  geom_hex(aes(mean, distance), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE) + ylim(c(-8,8)) + xlim(c(0,14)) + 
  ylab("Epsilon (Naive Batch1) - Epsilon (Naive Batch2)") + 
  xlab("Mean Expression log(Mu)") + theme_minimal(base_size = 12)

# Active cells
ggplot(data.frame(mean = log(rowMeans(matrix(c(colMedians(MCMC.batch1.active@parameters$mu),
                                               colMedians(MCMC.batch2.active@parameters$mu)),
                                             ncol = 2, 
                                             nrow = length(colMedians(MCMC.batch1.naive@parameters$mu)),
                                             byrow = FALSE))),
  distance = colMedians(MCMC.batch1.active@parameters$epsilon) -
    colMedians(MCMC.batch2.active@parameters$epsilon))) +
  geom_hex(aes(mean, distance), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE) + ylim(c(-8,8)) + xlim(c(0,14)) +
  ylab("Epsilon (Active Batch1) - Epsilon (Active Batch2)") + 
  xlab("Mean Expression log(Mu)") + theme_minimal(base_size = 12)

ggplot(data.frame(eps1 = colMedians(MCMC.batch1.active@parameters$epsilon),
                  eps2 = colMedians(MCMC.batch2.active@parameters$epsilon))) +
  geom_hex(aes(eps1, eps2), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE) 


# Differences between distances
ggplot(data.frame(mean = Test1$TableResDisp$MeanOverall,
  distance = Test1$TableResDisp$ResDispDistance - Test2$TableResDisp$ResDispDistance)) +
  geom_hex(aes(log(mean), distance), bins = 100) +
  scale_fill_gradientn("", colours = colorRampPalette(c("dark blue", "yellow", "dark red"))(100), guide=FALSE) + ylim(c(-8,8)) + xlim(c(0,14))

```

## Figure S5 C+D

Biological interpretation of structured variability

The purpose of this analysis is to guide the biological interpretation when outlying 
expression values are present.
We mixed 5 activated cells into a population of naive cells and compare variability
between the mixed population and the naive population.

```{r}
# Read in data
MCMC.naive <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Mixed_Tcells/MCMC_naive.rds")
MCMC.mixed <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Mixed_Tcells/MCMC_mixed.rds")
Data.naive <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Mixed_Tcells/Data_naive.rds")
Data.mixed <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Revisions/Mixed_Tcells/Data_mixed.rds")

# Exclude lowly expressed genes
genes <- colMedians(MCMC.naive@parameters$mu) > 1 & colMedians(MCMC.mixed@parameters$mu) > 1

Test <- BASiCS_TestDE(MCMC.naive, MCMC.mixed, EpsilonM = 1,
                      Plot = FALSE, PlotOffset = FALSE,
                      GroupLabel1 = "Naive", GroupLabel2 = "Mixed", 
                      GenesSelect = genes)

# Find genes more variable and more expressed in mixed condition
genes.DV <- Test$TableResDisp$GeneName[Test$TableResDisp$ResultDiffResDisp == "Mixed+" &
                                         Test$TableMean$ResultDiffMean == "Mixed+"]
mouse.genes[genes.DV,2]

# Normalize data
# Offset correction
OffSet <- function(MCMC1, MCMC2){
  OS <- median(rowSums(MCMC1@parameters$mu)/rowSums(MCMC2@parameters$mu))
  OS
}

OS <- OffSet(MCMC.mixed, MCMC.naive)

MCMC.mixed@parameters$mu <- MCMC.mixed@parameters$mu / OS
MCMC.mixed@parameters$phi <- MCMC.mixed@parameters$phi * OS

norm.mixed <- BASiCS_DenoisedCounts(Data.mixed, MCMC.mixed)
norm.naive <- BASiCS_DenoisedCounts(Data.naive, MCMC.naive)

# Visualize the expression of Il2
genes <- c("ENSMUSG00000027720")

df <- data.frame(value = c(as.vector(t(norm.mixed[genes,])), as.vector(t(norm.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(norm.mixed)),
                                  rep(genes, each = ncol(norm.naive)))),
                 Condition = factor(c(rep("Mixed", length(as.vector(t(norm.mixed[genes,])))),
                                      rep("Naive", length(as.vector(t(norm.naive[genes,]))))), 
                                    levels = c("Mixed", "Naive"))
                 )
final.jitter <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition), size = 2) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + 
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))

# Plot heatmap of DV genes
pheatmap(log10(norm.mixed[genes.DV,] + 1), cluster_cols = FALSE, color = viridis(100),
         border_color = NA, cellwidth = 8, cellheight = 8, fontsize = 8, show_colnames = FALSE,
         labels_row = mouse.genes[genes.DV,2], 
         annotation_col = data.frame(row.names = colnames(norm.mixed),
                                     activation_state = sapply(colnames(norm.mixed),
                                      function(n){unlist(strsplit(n, "\\."))[4]})))
```

```{r}
# Save final figure
final <- plot_grid(Diff.LogFC, NULL, final.jitter, NULL, ncol = 2, nrow = 2, labels = "AUTO")

ggsave(filename = "../../Analysis/Manuscript/Figures/Revisions/FigS3.pdf", 
       final, width = 12, height = 12)
```