---
title: "CD4 T cell differentiation"
author: "Nils Eling"
date: "16/04/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script repsoduces Figure 6 and Figure S6

This script reproduces the analysis doen on Loennberg et al.'s data.
Here, wer want to detect changes in varibility during immune differentation.

# Figure 6

```{r}
# Load the libraries 
library(BASiCS)
library(cowplot)
library(openxlsx)
library(RColorBrewer)

mouse.genes <- read.table("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Data/Genenames.txt", sep = '\t', header = TRUE, stringsAsFactors = FALSE)
rownames(mouse.genes) <- mouse.genes[,1]
```

```{r}
# Read in MCMC chains run on cells from day 2, day 4 and day7 post-infection
MCMCs <- readRDS("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/All_CD4diff.rds")

# Perform two sets of tests:
# 1. Testing on log2FC in mean expression > 1
# 2. Testing on log2FC in mean expression > 0 to find genes that don't change
#    in mean expression.
Test.day2_vs_day4 <- BASiCS_TestDE(Chain1 = MCMCs$day4, Chain2 = MCMCs$day2, 
                                   EpsilonM = 1, OffSet = TRUE, 
                                   GroupLabel2 = "Day2", GroupLabel1 = "Day4",
                                   Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName")

Test.day2_vs_day4.logFC0 <- BASiCS_TestDE(Chain1 = MCMCs$day4, Chain2 = MCMCs$day2, 
                                          EpsilonM = 0, OffSet = TRUE,
                                          GroupLabel2 = "Day2", GroupLabel1 = "Day4",
                                          Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName")

Test.day4_vs_day7 <- BASiCS_TestDE(Chain1 = MCMCs$day7, Chain2 = MCMCs$day4, 
                                   EpsilonM = 1, OffSet = TRUE, 
                                   GroupLabel2 = "Day4", GroupLabel1 = "Day7",
                                   Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName")

Test.day4_vs_day7.logFC0 <- BASiCS_TestDE(Chain1 = MCMCs$day7, Chain2 = MCMCs$day4, 
                                          EpsilonM = 0, OffSet = TRUE,
                                          GroupLabel2 = "Day4", GroupLabel1 = "Day7",
                                          Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName")

Test.day2_vs_day7 <- BASiCS_TestDE(Chain1 = MCMCs$day7, Chain2 = MCMCs$day2, 
                                   EpsilonM = 1, OffSet = TRUE, 
                                   GroupLabel2 = "Day2", GroupLabel1 = "Day7",
                                   Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName")

Test.day2_vs_day7.logFC0 <- BASiCS_TestDE(Chain1 = MCMCs$day7, Chain2 = MCMCs$day2, 
                                   EpsilonM = 0, OffSet = TRUE, 
                                   GroupLabel2 = "Day2", GroupLabel1 = "Day7",
                                   Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName")


# Collect genes with statistically significant changes in variability between all comparisons
Sig.2_4 <- Test.day2_vs_day4$TableResDisp[Test.day2_vs_day4$TableResDisp$ResultDiffResDisp != "NoDiff",]
Sig.2_4 <- Sig.2_4[order(Sig.2_4$ResDispDistance, decreasing = TRUE),]
Sig.2_4$Names <- mouse.genes[Sig.2_4$GeneName,2]

write.csv(Sig.2_4, "/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/CD4diff/DV_day2_day4.csv")

Sig.2_7 <- Test.day2_vs_day7$TableResDisp[Test.day2_vs_day7$TableResDisp$ResultDiffResDisp != "NoDiff",]
Sig.2_7 <- Sig.2_7[order(Sig.2_7$ResDispDistance, decreasing = TRUE),]
Sig.2_7$Names <- mouse.genes[Sig.2_7$GeneName,2]

write.csv(Sig.2_7, "/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/CD4diff/DV_day2_day7.csv")

Sig.4_7 <- Test.day4_vs_day7$TableResDisp[Test.day4_vs_day7$TableResDisp$ResultDiffResDisp != "NoDiff",]
Sig.4_7 <- Sig.4_7[order(Sig.4_7$ResDispDistance, decreasing = TRUE),]
Sig.4_7$Names <- mouse.genes[Sig.4_7$GeneName,2]

write.csv(Sig.4_7, "/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/CD4diff/DV_day4_day7.csv")
```

## Figure 6 A

First, we will examine global changes in over-dispersion for genes that 
don't change in mean expression.

```{r}
# Find genes that are not DE in any condition
genes <- intersect(intersect(Test.day2_vs_day4.logFC0$TableDisp$GeneName, 
                   Test.day2_vs_day7.logFC0$TableDisp$GeneName),
                   Test.day4_vs_day7.logFC0$TableDisp$GeneName)

# Look at delta values for these genes
df <- data.frame(delta = c(Test.day2_vs_day4.logFC0$TableDisp$Disp2[
                             match(genes, Test.day2_vs_day4.logFC0$TableDisp$GeneName)
                             ],
                           Test.day2_vs_day4.logFC0$TableDisp$Disp1[
                             match(genes, Test.day2_vs_day4.logFC0$TableDisp$GeneName)
                             ],
                           Test.day4_vs_day7.logFC0$TableDisp$Disp1[
                             match(genes, Test.day4_vs_day7.logFC0$TableDisp$GeneName)
                           ]),
                 Day = factor(c(rep("Day 2", length(genes)), rep("Day 4", length(genes)), 
                         rep("Day 7", length(genes))), levels = c("Day 2", "Day 4", "Day 7")))

# Plot boxplots
bp.delta <- ggplot(df) + geom_boxplot(aes(Day, log(delta), fill = Day), outlier.shape = NA) +
  scale_fill_manual(values = c("#fdd49e","#ef6548" ,"#7f0000")) + theme(legend.position = 'none')
```

## Figure 6 B

Now we will search for genes that show continous changes in variability 
or different patterns

```{r}
# Remove genes that are lowly expressed in all conditions from testing
genes <- !(is.na(MCMCs$day2@parameters$epsilon[1,]) | is.na(MCMCs$day4@parameters$epsilon[1,]) |
            is.na(MCMCs$day7@parameters$epsilon[1,]))

Test.day2_vs_day4.sets <- BASiCS_TestDE(Chain1 = MCMCs$day4, Chain2 = MCMCs$day2, 
                                   EpsilonM = 1, OffSet = TRUE, 
                                   GroupLabel2 = "Day2", GroupLabel1 = "Day4",
                                   Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName",
                                   GenesSelect = genes)

Test.day4_vs_day7.sets <- BASiCS_TestDE(Chain1 = MCMCs$day7, Chain2 = MCMCs$day4, 
                                   EpsilonM = 1, OffSet = TRUE, 
                                   GroupLabel2 = "Day4", GroupLabel1 = "Day7",
                                   Plot = FALSE, PlotOffset = FALSE, OrderVariable = "GeneName",
                                   GenesSelect = genes)

# 1st: up-up
genenames <- intersect(
  Test.day2_vs_day4.sets$TableResDisp$GeneName[Test.day2_vs_day4.sets$TableResDisp$ResultDiffResDisp == "Day4+"],
  Test.day4_vs_day7.sets$TableResDisp$GeneName[Test.day4_vs_day7.sets$TableResDisp$ResultDiffResDisp == "Day7+"]
)

cur_df <- data.frame(epsilon = c(Test.day2_vs_day4.sets$TableResDisp$ResDisp2[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day2_vs_day4.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day4_vs_day7.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day4_vs_day7.sets$TableResDisp$GeneName )
                           ]),
                 Day = factor(c(rep("Day 2", length(genenames)), rep("Day 4", length(genenames)), 
                         rep("Day 7", length(genenames))), levels = c("Day 2", "Day 4", "Day 7")),
                 Genes = rep(genenames, 3)) 

plot.1 <- ggplot(cur_df, aes(group = Genes)) + 
  geom_line(aes(x = Day, y = epsilon, alpha = 0.3)) +
  theme(legend.position = 'none', axis.title.x = element_blank()) +
  scale_x_discrete(expand=c(0.05,0.05)) +
  annotate("text", x=2.5, y=4, label= paste(length(genenames), " genes", sep = ""), size = 5) 

write.table(genenames, 
            "Results/Differential_testing/CD4diff/Up-Up.txt", quote = FALSE, 
            col.names = FALSE, row.names = FALSE)

# 2nd: down-down
genenames <- intersect(
  Test.day2_vs_day4.sets$TableResDisp$GeneName[Test.day2_vs_day4.sets$TableResDisp$ResultDiffResDisp == "Day2+"],
  Test.day4_vs_day7.sets$TableResDisp$GeneName[Test.day4_vs_day7.sets$TableResDisp$ResultDiffResDisp == "Day4+"]
)

cur_df <- data.frame(epsilon = c(Test.day2_vs_day4.sets$TableResDisp$ResDisp2[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day2_vs_day4.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day4_vs_day7.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day4_vs_day7.sets$TableResDisp$GeneName )
                           ]),
                 Day = factor(c(rep("Day 2", length(genenames)), rep("Day 4", length(genenames)), 
                         rep("Day 7", length(genenames))), levels = c("Day 2", "Day 4", "Day 7")),
                 Genes = rep(genenames, 3)) 

plot.2 <- ggplot(cur_df, aes(group = Genes)) + 
  geom_line(aes(x = Day, y = epsilon, alpha = 0.3)) +
  theme(legend.position = 'none', axis.title.x = element_blank()) +
  scale_x_discrete(expand=c(0.05,0.05)) +
  annotate("text", x=2.5, y=4, label= paste(length(genenames), " genes", sep = ""), size = 5) 

write.table(genenames, 
            "/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/CD4diff/Down-Down.txt", quote = FALSE, 
            col.names = FALSE, row.names = FALSE)

# 3rd: down-up
genenames <- intersect(
  Test.day2_vs_day4.sets$TableResDisp$GeneName[Test.day2_vs_day4.sets$TableResDisp$ResultDiffResDisp == "Day2+"],
  Test.day4_vs_day7.sets$TableResDisp$GeneName[Test.day4_vs_day7.sets$TableResDisp$ResultDiffResDisp == "Day7+"]
)

cur_df <- data.frame(epsilon = c(Test.day2_vs_day4.sets$TableResDisp$ResDisp2[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day2_vs_day4.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day4_vs_day7.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day4_vs_day7.sets$TableResDisp$GeneName )
                           ]),
                 Day = factor(c(rep("Day 2", length(genenames)), rep("Day 4", length(genenames)), 
                         rep("Day 7", length(genenames))), levels = c("Day 2", "Day 4", "Day 7")),
                 Genes = rep(genenames, 3)) 

plot.3 <- ggplot(cur_df, aes(group = Genes)) + 
  geom_line(aes(x = Day, y = epsilon, alpha = 0.3)) +
  theme(legend.position = 'none', axis.title.x = element_blank()) +
  scale_x_discrete(expand=c(0.05,0.05)) +
  annotate("text", x=2.5, y=4, label= paste(length(genenames), " genes", sep = ""), size = 5) 

write.table(genenames, 
            "/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/CD4diff/Down-Up.txt", quote = FALSE, 
            col.names = FALSE, row.names = FALSE)

# 4th: up-down
genenames <- intersect(
  Test.day2_vs_day4.sets$TableResDisp$GeneName[Test.day2_vs_day4.sets$TableResDisp$ResultDiffResDisp == "Day4+"],
  Test.day4_vs_day7.sets$TableResDisp$GeneName[Test.day4_vs_day7.sets$TableResDisp$ResultDiffResDisp == "Day4+"]
)

cur_df <- data.frame(epsilon = c(Test.day2_vs_day4.sets$TableResDisp$ResDisp2[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day2_vs_day4.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day2_vs_day4.sets$TableResDisp$GeneName)
                           ],
                           Test.day4_vs_day7.sets$TableResDisp$ResDisp1[
                             match(genenames,Test.day4_vs_day7.sets$TableResDisp$GeneName )
                           ]),
                 Day = factor(c(rep("Day 2", length(genenames)), rep("Day 4", length(genenames)), 
                         rep("Day 7", length(genenames))), levels = c("Day 2", "Day 4", "Day 7")),
                 Genes = rep(genenames, 3)) 

plot.4 <- ggplot(cur_df, aes(group = Genes)) + 
  geom_line(aes(x = Day, y = epsilon, alpha = 0.3)) +
  theme(legend.position = 'none', axis.title.x = element_blank()) +
  scale_x_discrete(expand=c(0.05,0.05)) +
  annotate("text", x=2.5, y=4, label= paste(length(genenames), " genes", sep = ""), size = 5) 

write.table(genenames, 
            "/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Results/Differential_testing/CD4diff/Up-Down.txt", quote = FALSE, 
            col.names = FALSE, row.names = FALSE)

dist.plot <- plot_grid(plot.1, plot.2, plot.3, plot.4, ncol = 2, nrow = 2, align = 'hv')

```

## Figure 6 C

Now we visualize individual genes that change in expression variability between 
day 2 and day 4.

```{r}
# Normalize data
input <- read.table("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Data/Test_Data/CD4_diff.txt", sep = "\t")

#### Read in Spike-ins
ERCC.conc <- read.table("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on//Data/Raw_data/CD4_Diff/ERCC_malaria.txt", header=TRUE, sep = "\t", fill = TRUE)

ERCC.num <- matrix(data=NA, nrow=nrow(ERCC.conc), ncol=1)
ERCC.num[,1] <- (ERCC.conc[,2]*(10^(-18)))*(6.0221417*(10^23))
rownames(ERCC.num) <- ERCC.conc[,1]

SpikeInput <- ERCC.num[rownames(input)[grepl("ERCC", rownames(input))],1]
SpikeInput.1 <- data.frame("Name" = names(SpikeInput),
                           "Molecules" = SpikeInput,
                           stringsAsFactors = FALSE)

#### Create Data objects for each condition
cur_input <- input[,grepl("2_", colnames(input))]
cur_chips <- sapply(colnames(cur_input), function(n){unlist(strsplit(n, "\\."))[1]})
Data.Day2 <- newBASiCS_Data(Counts = cur_input, 
                            Tech = grepl("ERCC", rownames(input)), 
                            SpikeInfo = SpikeInput.1, 
                            BatchInfo = cur_chips)

cur_input <- input[,grepl("4_", colnames(input))]
cur_chips <- sapply(colnames(cur_input), function(n){unlist(strsplit(n, "\\."))[1]})
Data.Day4 <- newBASiCS_Data(Counts = cur_input, 
                            Tech = grepl("ERCC", rownames(input)), 
                            SpikeInfo = SpikeInput.1, 
                            BatchInfo = cur_chips)

# Offset correction
OffSet <- function(MCMC1, MCMC2){
  OS <- median(rowSums(MCMC1@parameters$mu)/rowSums(MCMC2@parameters$mu))
  OS
}

OS <- OffSet(MCMCs$day2, MCMCs$day4)

MCMCs$day2@parameters$mu <- MCMCs$day2@parameters$mu / OS
MCMCs$day2@parameters$phi <- MCMCs$day2@parameters$phi * OS

Counts.day2 <- BASiCS_DenoisedCounts(Data.Day2, MCMCs$day2)
Counts.day4 <- BASiCS_DenoisedCounts(Data.Day4, MCMCs$day4)

rownames(Counts.day2) <- mouse.genes[rownames(Counts.day2),2]
rownames(Counts.day4) <- mouse.genes[rownames(Counts.day4),2]

# First genes that increase in variability between day 2 and day 4 
# Cxcr5 - Day4
df.gene <- data.frame(value = c(log10(Counts.day2["Cxcr5",] + 1),
                                log10(Counts.day4["Cxcr5",] + 1)),
                      Day = factor(c(rep("Day 2", ncol(Counts.day2)), 
                                      rep("Day 4", ncol(Counts.day4))),
                                    levels= c("Day 2", "Day 4")))

jitter.Cxcr5 <- ggplot(df.gene, aes(Day, value)) + 
  geom_boxplot(aes(fill = Day), alpha = 0.3, outlier.shape = NA, width=0.25)+
  geom_jitter(aes(fill = Day), width=0.25, shape=21) + ylab("log10(Cxcr5 expr)") + xlab("") +
  scale_color_manual(values = c("#fdd49e", "#ef6548")) + scale_fill_manual(values = c("#fdd49e", "#ef6548")) +
  theme(legend.position=c(0,1), legend.justification=c(0,1))+ ylim(c(-0.1,6.5))

# Tigit - Day4
df.gene <- data.frame(value = c(log10(Counts.day2["Tigit",] + 1),
                                log10(Counts.day4["Tigit",] + 1)),
                      Day = factor(c(rep("Day 2", ncol(Counts.day2)), 
                                      rep("Day 4", ncol(Counts.day4))),
                                    levels= c("Day 2", "Day 4")))

jitter.Tigit <- ggplot(df.gene, aes(Day, value)) + 
  geom_boxplot(aes(fill = Day), alpha = 0.3, outlier.shape = NA, width=0.25)+
  geom_jitter(aes(fill = Day), width=0.25, shape=21) + ylab("log10(Tigit expr)") + xlab("") +
  scale_color_manual(values = c("#fdd49e", "#ef6548")) + scale_fill_manual(values = c("#fdd49e", "#ef6548")) +
  theme(legend.position='none', legend.justification=c(0,1))+ ylim(c(-0.1,6.5))

# Tyk2 - Day4
df.gene <- data.frame(value = c(log10(Counts.day2["Tyk2",] + 1),
                                log10(Counts.day4["Tyk2",] + 1)),
                      Day = factor(c(rep("Day 2", ncol(Counts.day2)), 
                                      rep("Day 4", ncol(Counts.day4))),
                                    levels= c("Day 2", "Day 4")))

jitter.Tyk2 <- ggplot(df.gene, aes(Day, value)) + 
  geom_boxplot(aes(fill = Day), alpha = 0.3, outlier.shape = NA, width=0.25)+
  geom_jitter(aes(fill = Day), width=0.25, shape=21) + ylab("log10(Tyk2 expr)") + xlab("") +
  scale_color_manual(values = c("#fdd49e", "#ef6548")) + scale_fill_manual(values = c("#fdd49e", "#ef6548")) +
  theme(legend.position='none', legend.justification=c(0,1))+ ylim(c(-0.1,6.5))

# Now genes that decrease in variability between day 2 and day 4
# Ikzf4 - Day 2
df.gene <- data.frame(value = c(log10(Counts.day2["Ikzf4",] + 1),
                                log10(Counts.day4["Ikzf4",] + 1)),
                      Day = factor(c(rep("Day 2", ncol(Counts.day2)), 
                                      rep("Day 4", ncol(Counts.day4))),
                                    levels= c("Day 2", "Day 4")))

jitter.Ikzf4 <- ggplot(df.gene, aes(Day, value)) + 
  geom_boxplot(aes(fill = Day), alpha = 0.3, outlier.shape = NA, width=0.25)+
  geom_jitter(aes(fill = Day), width=0.25, shape=21) + ylab("log10(Ikzf4 expr)") + xlab("") +
  scale_color_manual(values = c("#fdd49e", "#ef6548")) + scale_fill_manual(values = c("#fdd49e", "#ef6548")) +
  theme(legend.position='none', legend.justification=c(0,1))+ ylim(c(-0.1,6.5))

# Ly6c1 - Day 2
df.gene <- data.frame(value = c(log10(Counts.day2["Ly6c1",] + 1),
                                log10(Counts.day4["Ly6c1",] + 1)),
                      Day = factor(c(rep("Day 2", ncol(Counts.day2)), 
                                      rep("Day 4", ncol(Counts.day4))),
                                    levels= c("Day 2", "Day 4")))

jitter.Ly6c1 <- ggplot(df.gene, aes(Day, value)) + 
  geom_boxplot(aes(fill = Day), alpha = 0.3, outlier.shape = NA, width=0.25)+
  geom_jitter(aes(fill = Day), width=0.25, shape=21) + ylab("log10(Ly6c1 expr)") + xlab("") +
  scale_color_manual(values = c("#fdd49e", "#ef6548")) + scale_fill_manual(values = c("#fdd49e", "#ef6548")) +
  theme(legend.position='none', legend.justification=c(0,1))+ ylim(c(-0.1,6.5))

# Tbx21 - Day 2
df.gene <- data.frame(value = c(log10(Counts.day2["Tbx21",] + 1),
                                log10(Counts.day4["Tbx21",] + 1)),
                      Day = factor(c(rep("Day 2", ncol(Counts.day2)), 
                                      rep("Day 4", ncol(Counts.day4))),
                                    levels= c("Day 2", "Day 4")))

jitter.Tbx21 <- ggplot(df.gene, aes(Day, value)) + 
  geom_boxplot(aes(fill = Day), alpha = 0.3, outlier.shape = NA, width=0.25)+
  geom_jitter(aes(fill = Day), width=0.25, shape=21) + ylab("log10(Tbx21 expr)") + xlab("") +
  scale_color_manual(values = c("#fdd49e", "#ef6548")) + scale_fill_manual(values = c("#fdd49e", "#ef6548")) +
  theme(legend.position='none', legend.justification=c(0,1)) + ylim(c(-0.1,6.5))

# Ratio of Tbx21 cells
length(which(Counts.day2["Tbx21",] > 0))/ncol(Counts.day2)

jitter.final <- plot_grid(jitter.Cxcr5, jitter.Tigit, jitter.Tyk2, 
                          jitter.Ikzf4, jitter.Ly6c1, jitter.Tbx21,
                          align = "hv", nrow = 2, ncol = 3, labels = c("C", NA, NA, "D", NA, NA))

```

## Figure 6 D

We can visualize expression dynamics for Tbx21 and Cxcr5 changes in mean expression
and expression variability.

```{r}
genes <- c("Cxcr5", "Tbx21")
rownames(Test.day2_vs_day4$TableResDisp) <- mouse.genes[Test.day2_vs_day4$TableResDisp$GeneName,2]
rownames(Test.day2_vs_day4$TableMean) <-  mouse.genes[Test.day2_vs_day4$TableMean$GeneName,2]
rownames(Test.day4_vs_day7$TableResDisp) <- mouse.genes[Test.day4_vs_day7$TableResDisp$GeneName,2]
rownames(Test.day4_vs_day7$TableMean) <- mouse.genes[Test.day4_vs_day7$TableMean$GeneName,2]
rownames(Test.day2_vs_day7$TableResDisp) <- mouse.genes[Test.day2_vs_day7$TableResDisp$GeneName,2]
rownames(Test.day2_vs_day7$TableMean) <- mouse.genes[Test.day2_vs_day7$TableMean$GeneName,2]

df <- data.frame(mu = c(Test.day2_vs_day4$TableMean[genes,"Mean2"],
                        Test.day2_vs_day4$TableMean[genes,"Mean1"],
                        Test.day2_vs_day7$TableMean[genes,"Mean1"]),
                 epsilon = c(Test.day2_vs_day4$TableResDisp[genes,"ResDisp2"],
                        Test.day2_vs_day4$TableResDisp[genes,"ResDisp1"],
                        Test.day2_vs_day7$TableResDisp[genes,"ResDisp1"]),
                 Genes = factor(rep(genes, 3)),
                 Day = factor(rep(c("Day 2", "Day 4", "Day 7"), each = 2)))

df.arrows <- data.frame(x = df[1:4,1],
                        y = df[1:4,2],
                        xend = df[3:6,1],
                        yend = df[3:6,2],
                        Genes = rep(genes, 2))

trace.plot <- ggplot(df) + 
  scale_colour_manual(values = setNames(c("indianred4", "royalblue4"), 
                                        c("Cxcr5", "Tbx21"))) + 
  geom_segment(data=df.arrows, 
        aes(x=log(x), y=y, xend=log(xend), yend=yend, colour = Genes),
        lineend="round",arrow=arrow(length=unit(0.2, "inches")), size=2) + xlab("log(Expr)") + ylab("Epsilon \n Less variable <-> More variable")
```

Save final figure

```{r}
# Final figure
Fig6 <- plot_grid(bp.delta, dist.plot, jitter.final, trace.plot, ncol = 2, nrow = 2,
                   labels = c("A", "B", NA, "D"))

ggsave("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis/Manuscript/Figures/Fig6.pdf", plot = Fig6, width = 10, height = 10)

```

# Figure S6

This script aims to dissect the differentiation process of CD4 T cells in more detail.
We will use genes annotated in Lönnberg et al.

## Figure S6A

We will use the gene modules detected by GPfates as displayed in Figure 3E in
Lönnberg et al.

```{r}
Th1 <- c("Cxcr6", "Gna15", "S100a4", "Gzmb", "Chsy1", "Ccr2", "Ccr5", "Ifngr1", "Sdcbp2",
         "Gp49a", "Plek", "Lgals3", "Havcr2", "Sytl2", "F2rl2", "Nkg7", "Il18rap", "Ctsw",
         "Entpd1", "Cd38", "Myo1f", "F2r", "Xdh", "Lilrb4", "Ahnak", "S100a6", "Serpinb6b",
         "Pglyrp1", "Ccl5", "Tmprss13", "Il12rb2", "Sla2", "Il18r1", "Ifng", "Ctsd",
         "S1pr1", "Tmem154", "Capn2", "Syt11", "Mfsd10", "Cd44", "Sema4a", "Ccl4",
         "Emp3", "4930453N24Rik", "Plac8", "Prex1", "Anxa2", "AU020206", "Glrx")

Tfh <- c("Asap1", "Tcf7", "Rgs10", "Tbc1d4", "Gm10275", "Zfp36l1", "Cxcr5",
         "Ddx24", "Rps12-ps3", "Rps20", "Erdr1", "Tnfsf8", "Cdk2ap2", "Kcnn4",
         "Rpl10a", "D19Bwg1357e", "Ivns1abp", "Bcas2", "Znf512b", "Tagap1",
         "Ptma", "Nop10", "Tnfrsf26", "Alyref", "Rpl36a", "Npm1", "Dennd2d", "Lsm7",
         "Il16", "Trp53", "Clec2i", "Rpl10", "Sarnp",
         "Srsf2", "Inpp4b", "Rps2", "Emg1", "Phb", "Hnrnpr", "Nap1l1", "Cops6",
         "Prdx6", "Lsm3", "Adk", "Mrps6", "Cdk4")

genes.ENS.Th1 <- mouse.genes[match(Th1, mouse.genes[,2]),1]
genes.ENS.Tfh <- mouse.genes[match(Tfh, mouse.genes[,2]),1]

genes.ENS.Th1 <- genes.ENS.Th1[!is.na(match(genes.ENS.Th1, Test.day2_vs_day4$TableMean$GeneName))]
genes.ENS.Tfh <- genes.ENS.Tfh[!is.na(match(genes.ENS.Tfh, Test.day2_vs_day4$TableMean$GeneName))]

final.genes <- c(genes.ENS.Th1, rev(genes.ENS.Tfh))

DE.2_4 <- ifelse(Test.day2_vs_day4$TableMean[match(final.genes,
                                Test.day2_vs_day4$TableMean$GeneName),"ResultDiffMean"] == "Day2+", "Down",
                 ifelse(Test.day2_vs_day4$TableMean[match(final.genes,
                                Test.day2_vs_day4$TableMean$GeneName),"ResultDiffMean"] == "Day4+", "Up",
                       "NoDiff"))
DE.4_7 <- ifelse(Test.day4_vs_day7$TableMean[match(final.genes,
                                Test.day4_vs_day7$TableMean$GeneName),"ResultDiffMean"] == "Day4+", "Down",
                 ifelse(Test.day4_vs_day7$TableMean[match(final.genes,
                                Test.day4_vs_day7$TableMean$GeneName),"ResultDiffMean"] == "Day7+", "Up",
                       "NoDiff"))

DV.2_4 <- ifelse(Test.day2_vs_day4$TableResDisp[match(final.genes,
                                Test.day2_vs_day4$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day2+", "Down",
                 ifelse(Test.day2_vs_day4$TableResDisp[match(final.genes,
                                Test.day2_vs_day4$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day4+", "Up",
                       "NoDiff"))
DV.4_7 <- ifelse(Test.day4_vs_day7$TableResDisp[match(final.genes,
                                Test.day4_vs_day7$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day4+", "Down",
                 ifelse(Test.day4_vs_day7$TableResDisp[match(final.genes,
                                Test.day4_vs_day7$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day7+", "Up",
                       "NoDiff"))

df <- data.frame(Day = rep(c("Day 2 vs Day 4", "Day 4 vs Day 7"), each = length(final.genes)),
                 Genes = factor(rep(mouse.genes[final.genes,2], 2), levels = rev(mouse.genes[final.genes,2])),
                 DE = c(DE.2_4, DE.4_7),
                 DV = c(DV.2_4, DV.4_7))


DE.module <- ggplot(df) + geom_point(aes(Day, Genes, colour = DE), size = 4) + coord_fixed(ratio = 1) + 
  scale_color_manual(values = c("blue", "grey", "red")) + theme(axis.text.x = element_text(angle = 90))

DV.module <- ggplot(df) + geom_point(aes(Day, Genes, colour = DV), size = 4) + coord_fixed(ratio = 1) +  
  scale_color_manual(values = c("green", "grey", "purple")) + theme(axis.text.x = element_text(angle = 90))
```

## Figure S6B

We will now take a closer look into the fate decision module

```{r}
# Important genes ordered by peak expression
stage <- c("Il2ra", "Tbx21", "Il2rb", "Cxcr5", "Cxcr3", "Il21", "Pdcd1", "Cxcr6", "Selplg",
            "Ccr2", "Id2", "Ifng", "Icos", "Ifngr1", "Ccr5")

stage.ENS <- mouse.genes[match(stage, mouse.genes[,2]),1]

stage.ENS <- stage.ENS[!is.na(match(stage.ENS, Test.day2_vs_day4$TableMean$GeneName))]

final.genes <- stage.ENS

DE.2_4 <- ifelse(Test.day2_vs_day4$TableMean[match(final.genes,
                                Test.day2_vs_day4$TableMean$GeneName),"ResultDiffMean"] == "Day2+", "Down",
                 ifelse(Test.day2_vs_day4$TableMean[match(final.genes,
                                Test.day2_vs_day4$TableMean$GeneName),"ResultDiffMean"] == "Day4+", "Up",
                       "NoDiff"))
DE.4_7 <- ifelse(Test.day4_vs_day7$TableMean[match(final.genes,
                                Test.day4_vs_day7$TableMean$GeneName),"ResultDiffMean"] == "Day4+", "Down",
                 ifelse(Test.day4_vs_day7$TableMean[match(final.genes,
                                Test.day4_vs_day7$TableMean$GeneName),"ResultDiffMean"] == "Day7+", "Up",
                       "NoDiff"))

DV.2_4 <- ifelse(Test.day2_vs_day4$TableResDisp[match(final.genes,
                                Test.day2_vs_day4$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day2+", "Down",
                 ifelse(Test.day2_vs_day4$TableResDisp[match(final.genes,
                                Test.day2_vs_day4$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day4+", "Up",
                       "NoDiff"))
DV.4_7 <- ifelse(Test.day4_vs_day7$TableResDisp[match(final.genes,
                                Test.day4_vs_day7$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day4+", "Down",
                 ifelse(Test.day4_vs_day7$TableResDisp[match(final.genes,
                                Test.day4_vs_day7$TableResDisp$GeneName),"ResultDiffResDisp"] == "Day7+", "Up",
                       "NoDiff"))

df <- data.frame(Day = rep(c("Day 2 vs Day 4", "Day 4 vs Day 7"), each = length(final.genes)),
                 Genes = factor(rep(mouse.genes[final.genes,2], 2), levels = rev(mouse.genes[final.genes,2])),
                 DE = c(DE.2_4, DE.4_7),
                 DV = c(DV.2_4, DV.4_7))


DE.selected <- ggplot(df) + geom_point(aes(Day, Genes, colour = DE), size = 4) + coord_fixed(ratio = 1) + 
  scale_color_manual(values = c("blue", "grey", "red")) + theme(axis.text.x = element_text(angle = 90))

DV.selected <- ggplot(df) + geom_point(aes(Day, Genes, colour = DV), size = 4) + coord_fixed(ratio = 1) +  
  scale_color_manual(values = c("green", "grey", "purple")) + theme(axis.text.x = element_text(angle = 90))

```

Cell cycle GO term

```{r}
Cell.cylce <- read.table("CD4_diff/human_experiments_ptms.tsv", sep = "\t", 
                         header = TRUE, stringsAsFactors = FALSE)
protein <- unique(Cell.cylce$gene)
mouse.human.prot <- read.table("CD4_diff/Human_Protein_to_Mouse_gene.txt", 
                               sep = "\t", header = TRUE, stringsAsFactors = FALSE)
genes <- mouse.human.prot[match(protein, mouse.human.prot$Human.protein.or.transcript.stable.ID)[!is.na(match(protein, mouse.human.prot$Human.protein.or.transcript.stable.ID))],1]

final.genes <- genes[!is.na(match(genes, Test.day2_vs_day4$TableMean$GeneName))]

df <- data.frame(Log2FCMean = Test.day2_vs_day4$TableMean$MeanLog2FC[
                          match(final.genes, Test.day2_vs_day4$TableMean$GeneName)],
                 DistanceEps = Test.day2_vs_day4$TableResDisp$ResDispDistance[
                          match(final.genes, Test.day2_vs_day4$TableResDisp$GeneName)],
                 DV = Test.day2_vs_day4$TableResDisp$ResultDiffResDisp[
                          match(final.genes, Test.day2_vs_day4$TableResDisp$GeneName)])

ggplot(df) + geom_point(aes(Log2FCMean, DistanceEps, colour = DV))

df <- data.frame(Log2FCMean = Test.day4_vs_day7$TableMean$MeanLog2FC[
                          match(final.genes, Test.day4_vs_day7$TableMean$GeneName)],
                 DistanceEps = Test.day4_vs_day7$TableResDisp$ResDispDistance[
                          match(final.genes, Test.day4_vs_day7$TableResDisp$GeneName)],
                 DV = Test.day4_vs_day7$TableResDisp$ResultDiffResDisp[
                          match(final.genes, Test.day4_vs_day7$TableResDisp$GeneName)])

ggplot(df) + geom_point(aes(Log2FCMean, DistanceEps, colour = DV))

```

Save these plots.

```{r}
FigS6 <- plot_grid(DE.module, DV.module, DE.selected, DV.selected, ncol = 4, nrow = 1)
ggsave("/Users/nils/Google Drive File Stream/My Drive/BASiCS_add-on/Analysis//Manuscript/Figures/Revisions/FigS6.pdf", FigS6, height = 18, width = 15)
```


