---
title: "Correction for Figure 5 and Figure S5"
author: "Nils Eling and Catalina Vallejos"
date: '`r Sys.Date()`'
output:
  html_document:
      code_folding: hide
      toc: true
      toc_float: 
        collapsed: false
      number_sections: false 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script reproduces the analysis performed for Figure 5 and Figure S5.

# Figure 5

First, we will load the genenames from biomart.

```{r}
library(biomaRt)

# Initialize mart
ensembl <- useMart("ensembl")

# Select dataset
ensembl <- useDataset(dataset = "mmusculus_gene_ensembl", 
                      mart = ensembl)

# Select gene ID and gene name
genenames <- getBM(attributes = c("ensembl_gene_id", "external_gene_name"),
                   mart = ensembl)
rownames(genenames) <- genenames$ensembl_gene_id
```

## Loading MCMC chains

Now, we will load the chains for the naive and activated CD4+ T cells.

```{r}
library(BASiCS)
MCMC_naive <- BASiCS_LoadChain("B6naive_cata", StoreDir = "../../MCMCs/")
MCMC_active <- BASiCS_LoadChain("B6active_cata", StoreDir = "../../MCMCs/")
```

## Convergence diagnostics

We use `BASiCS_diagPlot` in order to visualise the effective sample size (ESS)
obtained for gene-specific expression parameters. The ESS can be used to 
assess the mixing of the MCMC, with low ESS indicating poor mixing. 

```{r Convergence}
library(ggplot2)
library(cowplot)

p1 <- BASiCS_diagPlot(MCMC_naive)
p2 <- BASiCS_diagPlot(MCMC_active)
plot_grid(p1, p2)
```

The figures above suggest good mixing for most genes, although low ESS is 
observed for lowly expressed genes (minimum observed ESS is `r min(p1$data[,2])` 
and `r min(p2$data[,2])` within naive and active cells, respectively). 

To avoid making inferences for those genes, subsequent analyses will only focus
on those genes with ESS of at least 500. Below, we show more detailed diagnostic 
plots for those genes with the lowest ESS within each experimental condition.

```{r}
inclgenes <- (p1$data[,2] > 500) & (p2$data[,2] > 500) 

MCMC_naive_aux <- subset(MCMC_naive, 
                         Genes = rownames(MCMC_naive)[inclgenes])
MCMC_active_aux <- subset(MCMC_active, 
                          Genes = rownames(MCMC_active)[inclgenes])

p3 <- BASiCS_diagPlot(MCMC_naive_aux)
p4 <- BASiCS_diagPlot(MCMC_active_aux)
plot_grid(p3, p4)
```

After applying this inclusion criteria, `r sum(!inclgenes)` will be excluded (i.e.
`r sum(inclgenes)` will be considered within the differential test). 

To further assess the behaviour of the MCMc, we next provide more detailed 
diagnostic criteria for those genes with the lowest ESS within each group.

```{r, fig.width=10, fig.height=5}
plot(MCMC_naive_aux, Param = "mu", Gene = which(p3$data[,2] == min(p3$data[,2])))
plot(MCMC_active_aux, Param = "mu", Gene = which(p4$data[,2] == min(p4$data[,2])))
```

## Differential expression analysis

Subsequently, we apply the differential expression test. 

```{r}
# Perform differential testing between naive and activated CD4+ T cells
Test.NaiveActive <- BASiCS_TestDE(Chain1 = MCMC_active, 
                                  Chain2 = MCMC_naive, 
                                  EpsilonM = 1, 
                                  OffSet = TRUE, 
                                  GenesSelect = inclgenes,
                                  Plot = FALSE, 
                                  PlotOffset = FALSE, 
                                  EFDR_M = 0.1, EFDR_D = 0.1, EFDR_R = 0.1)

# Exlude the lowly expressed genes
Test.NaiveActive$TableResDisp <- Test.NaiveActive$TableResDisp[
  Test.NaiveActive$TableResDisp$ResultDiffResDisp != "ExcludedByUser" &
    Test.NaiveActive$TableResDisp$ResultDiffResDisp != "ExcludedFromTesting",]
Test.NaiveActive$TableMean <- Test.NaiveActive$TableMean[
  match(Test.NaiveActive$TableResDisp$GeneName, Test.NaiveActive$TableMean$GeneName),]
```

Generate DE and DV plot.

```{r}
# Check that tables are in the same order
table(Test.NaiveActive$TableMean$GeneName == Test.NaiveActive$TableResDisp$GeneName)
# Table below extended to include probabilities
df.DE <- data.frame(row.names = Test.NaiveActive$TableMean$GeneName,
                 DistEpsilon = Test.NaiveActive$TableResDisp$ResDispDistance,
                 ProbDiffEpsilon = Test.NaiveActive$TableResDisp$ProbDiffResDisp,
                 Log2FCmu = Test.NaiveActive$TableMean$MeanLog2FC,
                 ProbDiffMean = Test.NaiveActive$TableMean$ProbDiffMean,
                 Regulation = paste(Test.NaiveActive$TableMean$ResultDiffMean,
                               Test.NaiveActive$TableResDisp$ResultDiffResDisp, sep = "_"))
df.DE$Regulation <- ifelse(df.DE$Regulation == "Group1+_Group1+", 
                        "Up-regulated, higher variability", 
                   ifelse(df.DE$Regulation == "Group1+_Group2+", 
                          "Up-regulated, lower variability",
                          ifelse(df.DE$Regulation == "Group2+_Group2+", 
                                 "Down-regulated, lower variability",
                                 ifelse(df.DE$Regulation == "Group2+_Group1+", 
                                        "Down-regulated, higher variability", "Other"))))

df.DE$Genename <- genenames[rownames(df.DE),"external_gene_name"]
df.DE$GeneID <- rownames(df.DE)

# Plot the figure
fourGroups <- ggplot(df.DE, aes(Log2FCmu, DistEpsilon)) + geom_point(aes(colour=Regulation)) + 
  xlab("Log2FC mu") + ylab("Distance epsilon") + 
  scale_colour_manual(values = c("violetred4", "violetred2", 
                                 "grey90", "seagreen2", "seagreen4")) 
fourGroups

ggsave(file.path("~/Dropbox (Cambridge University)/BASiCSplus/Comparisons/CellSystems/",
                 "New_Figures_Results/Figure_5/Panel_A_cata.pdf"), 
       fourGroups, width = 12, height = 7)
```

Save Table S2.

```{r}
df.out <- df.DE[df.DE$Regulation != "Other",]
df.out <- df.out[order(df.out$Regulation),]
df.out$GeneID <- NULL

library(openxlsx)
write.xlsx(df.out, 
           file.path("~/Dropbox (Cambridge University)/BASiCSplus/Comparisons/CellSystems/",
                     "New_tables/TableS2_cata.xlsx"),
           row.names = TRUE)
```

Save gene names for enrichment analysis.

```{r}
# Write out genes to use with DAVID
write.table(as.data.frame(rownames(df.DE)), 
            "../../New_Figures_Results/Figure_5/background_cata.txt", sep = "\t", 
            row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(as.data.frame(rownames(df.DE)[df.DE$Regulation == "Up-regulated, lower variability"]),
            "../../New_Figures_Results/Figure_5/CD4_activation_DEup_DVdown_cata.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(as.data.frame(rownames(df.DE)[df.DE$Regulation == "Up-regulated, higher variability"]),
            "../../New_Figures_Results/Figure_5/CD4_activation_DEup_DVup_cata.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(as.data.frame(rownames(df.DE)[df.DE$Regulation == "Down-regulated, lower variability"]),
            "../../New_Figures_Results/Figure_5/CD4_activation_DEdown_DVdown_cata.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(as.data.frame(rownames(df.DE)[df.DE$Regulation == "Down-regulated, higher variability"]),
            "../../New_Figures_Results/Figure_5/CD4_activation_DEdown_DVup_cata.txt", 
            sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

## Visualisation of example genes

Normalize the counts.

```{r}
# Function for correcting differences in population wide RNA content
OffSetCorrection <- function(MCMC1, MCMC2){
  median(rowSums(MCMC1@parameters$mu)/rowSums(MCMC2@parameters$mu)) 
}
# Offset correction
Offset <- OffSetCorrection(MCMC_active, MCMC_naive)
Offset

cur_active <- MCMC_active
cur_active@parameters$mu <- MCMC_active@parameters$mu / Offset
cur_active@parameters$phi <- MCMC_active@parameters$phi * Offset

Data.NaiveB6 <- readRDS("../../Data/CD4_naive.rds")
Data.ActiveB6 <- readRDS("../../Data/CD4_active.rds")
Counts.naive <- BASiCS_DenoisedCounts(Data.NaiveB6, MCMC_naive)
Counts.active <- BASiCS_DenoisedCounts(Data.ActiveB6, cur_active)
rownames(Counts.active) <- genenames[rownames(Counts.active),2]
rownames(Counts.naive) <- genenames[rownames(Counts.naive),2]
```

Plot boxplots

Figure 5B, using the original genes

```{r}
genes <- c("Ncl", "Polr2l", "Cpsf2", "Sf3a3", "Plrg1") 
df <- data.frame(value = c(as.vector(t(Counts.active[genes,])),
                           as.vector(t(Counts.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(Counts.active)),
                                  rep(genes, each = ncol(Counts.naive))), 
                                levels = rev(genes)),
                 Condition = factor(c(rep("Active",
                                          length(as.vector(t(Counts.active[genes,])))),
                                      rep("Naive", 
                                          length(as.vector(t(Counts.naive[genes,]))))), 
                                    levels = c("Active", "Naive")))
final.lessVar <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition)) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + ylim(c(0,4)) +
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), 
        axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
final.lessVar

# Effect sizes and posterior probabilities
df.DE[df.DE$Genename %in% genes,]
```

As it can be seen above, some genes are no longer significant for either
the mean or the residual overdispersion test. Note that for *Ncl*, the 
probability of differential residual over-dispersion is high
(`r df.DE$ProbDiffEpsilon[which(df.DE$Genename == "Ncl")]`), but below the 
probability cut-off (`r Test.NaiveActive$DiffResDispSummary$ProbThreshold`). 

Replacement genes were proposed by Nils (see below). As *Polr1d* and *Naf1*
are mentioned in the text, 

```{r}
genes <- c("Larp1", "Snrpd2", "Polr2l", "Prpf31") 
df <- data.frame(value = c(as.vector(t(Counts.active[genes,])),
                           as.vector(t(Counts.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(Counts.active)),
                                  rep(genes, each = ncol(Counts.naive))), 
                                levels = rev(genes)),
                 Condition = factor(c(rep("Active",
                                          length(as.vector(t(Counts.active[genes,])))),
                                      rep("Naive", 
                                          length(as.vector(t(Counts.naive[genes,]))))), 
                                    levels = c("Active", "Naive")))
final.lessVar <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition)) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + ylim(c(0,4)) +
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), 
        axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
final.lessVar

ggsave("../../New_Figures_Results/Figure_5/Panel_B.pdf",final.lessVar, width = 5, height = 5)

# Effect sizes and posterior probabilities
df.DE[df.DE$Genename %in% genes,]
```

Figure 5C.

```{r}
genes <- c("Smad3", "Pou2f2", "Il2", "Cd274")
df <- data.frame(value = c(as.vector(t(Counts.active[genes,])), as.vector(t(Counts.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(Counts.active)),
                                  rep(genes, each = ncol(Counts.naive))), 
                                levels = rev(genes)),
                 Condition = factor(c(rep("Active",
                                          length(as.vector(t(Counts.active[genes,])))),
                                      rep("Naive", 
                                          length(as.vector(t(Counts.naive[genes,]))))), 
                                    levels = c("Active", "Naive")))
final.moreVar <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition)) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + ylim(c(0,4)) +
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
final.moreVar

ggsave("../../New_Figures_Results/Figure_5/Panel_C.pdf",final.moreVar, width = 5, height = 5)

# Effect sizes and posterior probabilities
df.DE[df.DE$Genename %in% genes,]
```

# Figure S5

## Figure S5 A+B

Comparison of DV genes detected by regression and non-regression model - 
focusing on the same genes that were included in the original analysis. 

```{r}
# Offset correction not required as applied before
inclgenes2 <- (colMedians(MCMC_active@parameters$mu) > 1 &
                 colMedians(MCMC_naive@parameters$mu) > 1)

library(VennDiagram)

# Load extra data
# Load chains of active and naive T cells generated by regression and 
# non-regression case
MCMC.naive.new <- readRDS("../../MCMCs/chain_B6naive_cata.rds")
MCMC.naive.old <- readRDS("../../MCMCs/chain_B6naive_noReg.rds")
MCMC.active.new <- readRDS("../../MCMCs/chain_B6active_cata.rds")
MCMC.active.old <- readRDS("../../MCMCs/chain_B6active_noReg.Rds")

# Perform testing between naive and active populations for regression and non-regression case
# First exclude genes same genes as before

# Perform testing on log2 FC = 2 as in Martinez-Jimenez et al, 2017 to detect changes in mean expression
Test_Regression.DE <- BASiCS_TestDE(MCMC.naive.new, MCMC.active.new, EpsilonM = 2,
                                    Plot = FALSE, PlotOffset = FALSE,
                                    GroupLabel1 = "Naive_Regression", 
                                    GroupLabel2 = "Active_Regression",
                                    GenesSelect = inclgenes2,
                                    EFDR_M = 0.1, EFDR_D = 0.1, EFDR_R = 0.1)

Test_NonRegression.DE <- BASiCS_TestDE(MCMC.naive.old, MCMC.active.old, EpsilonM = 2,
                                       Plot = FALSE, PlotOffset = FALSE,
                                       GroupLabel1 = "Naive_NonRegression", 
                                       GroupLabel2 = "Active_NonRegression",
                                       GenesSelect = inclgenes2,
                                       EFDR_M = 0.1, EFDR_D = 0.1, EFDR_R = 0.1)

# Perform testing on log2 FC = 0 as in Martinez-Jimenez et al, 2017 to detect changes in varaibility
Test_Regression.DV <- BASiCS_TestDE(MCMC.naive.new, MCMC.active.new, EpsilonM = 0,
                                    Plot = FALSE, PlotOffset = FALSE,
                                    GroupLabel1 = "Naive_Regression", 
                                    GroupLabel2 = "Active_Regression",
                                    GenesSelect = inclgenes2,
                                    EFDR_M = 0.1, EFDR_D = 0.1, EFDR_R = 0.1)

Test_NonRegression.DV <- BASiCS_TestDE(MCMC.naive.old, MCMC.active.old, EpsilonM = 0,
                                       Plot = FALSE, PlotOffset = FALSE,
                                       GroupLabel1 = "Naive_NonRegression", 
                                       GroupLabel2 = "Active_NonRegression",
                                       GenesSelect = inclgenes2,
                                       EFDR_M = 0.1, EFDR_D = 0.1, EFDR_R = 0.1)

# Figure S5A
# First we compare the differential mean expression test between the two models
# Genes that are detected as differentially expressed
NumberDE.Regression <- sum(Test_Regression.DE$TableMean$ResultDiffMean != "NoDiff" &
                             Test_Regression.DE$TableMean$ResultDiffMean != "ExcludedByUser")
NumberDE.NonRegression <- sum(Test_NonRegression.DE$TableMean$ResultDiffMean != "NoDiff" &
                                Test_NonRegression.DE$TableMean$ResultDiffMean != "ExcludedByUser")
Overlap <- length(intersect(Test_Regression.DE$TableMean$GeneName[
                                  Test_Regression.DE$TableMean$ResultDiffMean != "NoDiff" &
                                    Test_Regression.DE$TableMean$ResultDiffMean != "ExcludedByUser"],
                            Test_NonRegression.DE$TableMean$GeneName[
                                  Test_NonRegression.DE$TableMean$ResultDiffMean != "NoDiff" &
                                    Test_NonRegression.DE$TableMean$ResultDiffMean != "ExcludedByUser"]
                         ))
pdf("../../New_Figures_Results/Figure_S5/Fig_S5A_cata.pdf", width = 7, height = 5)
draw.pairwise.venn(area1 = NumberDE.NonRegression, area2 = NumberDE.Regression, cross.area = Overlap, 
                   category = c("Differential mean: Non Regression case", 
                                "Differential mean: Regression case"), euler.d = TRUE, scaled = TRUE,
                   fill = c("steelblue4", "coral3"))
dev.off()

# Figure S5B
# Genes that are not detected as differentially expressed
NumberDE.Regression <- sum(Test_Regression.DV$TableMean$ResultDiffMean == "NoDiff")
NumberDE.NonRegression <- sum(Test_NonRegression.DV$TableMean$ResultDiffMean == "NoDiff")
Overlap <- length(intersect(Test_Regression.DV$TableMean$GeneName[
                                  Test_Regression.DV$TableMean$ResultDiffMean == "NoDiff"],
                            Test_NonRegression.DV$TableMean$GeneName[
                                  Test_NonRegression.DV$TableMean$ResultDiffMean == "NoDiff"]
                         ))

pdf("../../New_Figures_Results/Figure_S5/Fig_S5ABi_cata.pdf", width = 7, height = 5)
draw.pairwise.venn(area1 = NumberDE.NonRegression, area2 = NumberDE.Regression, cross.area = Overlap, 
                   category = c("No differential mean: Non Regression case", 
                                "No differential mean: Regression case"), euler.d = TRUE, scaled = TRUE,
                   fill = c("steelblue4", "coral3"))
dev.off()

# Second, we compare the DV genes between regression and non-regression case using the deltas
# Select the genes that are not differentially expressed in both comparisons
genes <- intersect(Test_Regression.DV$TableMean$GeneName[
                                  Test_Regression.DV$TableMean$ResultDiffMean == "NoDiff"],
                            Test_NonRegression.DV$TableMean$GeneName[
                                  Test_NonRegression.DV$TableMean$ResultDiffMean == "NoDiff"]
                         )
NumberDV.Regression <- sum(Test_Regression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                         Test_Regression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_Regression.DV$TableDisp$GeneName %in% genes)
NumberDV.NonRegression <- sum(Test_NonRegression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                         Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_NonRegression.DV$TableDisp$GeneName %in% genes)
Overlap <- length(intersect(Test_Regression.DV$TableDisp$GeneName[
                                  Test_Regression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                                  Test_Regression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_Regression.DV$TableDisp$GeneName %in% genes
                                  ],
                            Test_NonRegression.DV$TableDisp$GeneName[
                                  Test_NonRegression.DV$TableDisp$ResultDiffDisp != "NoDiff" &
                                  Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_NonRegression.DV$TableDisp$GeneName %in% genes
                                  ]
                         ))

# Jaccard index
Overlap/(NumberDV.Regression + NumberDV.NonRegression - Overlap)

pdf("../../New_Figures_Results/Figure_S5/Fig_S5ABii_cata.pdf", width = 7, height = 5)
draw.pairwise.venn(area1 = NumberDV.NonRegression, area2 = NumberDV.Regression, cross.area = Overlap, 
                   category = c("Differential dispersion :Non Regression case", 
                                "Differential dispersion: Regression case"), euler.d = TRUE, scaled = TRUE,
                   fill = c("steelblue4", "coral3"))
dev.off()

# Number of genes that can be tested for: 1. differential residual over-dispersion
#                                         2. differential over-dispersion
sum(Test_Regression.DV$TableResDisp$ResultDiffResDisp != "ExcludedFromTesting" & 
                           Test_Regression.DV$TableResDisp$ResultDiffResDisp != "ExcludedByUser")
# 6944
sum(Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedFromTesting" &
                           Test_NonRegression.DV$TableDisp$ResultDiffDisp != "ExcludedByUser")
# 2758
```

## Figure S5 C+D

Biological interpretation of structured variability

The purpose of this analysis is to guide the biological interpretation when outlying 
expression values are present.
We mixed 5 activated cells into a population of naive cells and compare variability
between the mixed population and the naive population.

```{r}
# Read in data
MCMC.naive <- BASiCS_LoadChain("B6naive", StoreDir = "../../MCMCs/")
MCMC.mixed <- BASiCS_LoadChain("B6mixed", StoreDir = "../../MCMCs/")
Data.naive <- readRDS("../../Data/CD4_naive.rds")
Data.mixed <- readRDS("../../Data/CD4_mixed.rds")
```

```{r}
Test <- BASiCS_TestDE(MCMC.naive, MCMC.mixed, 
                      EpsilonM = 1,
                      Plot = FALSE, PlotOffset = FALSE,
                      GroupLabel1 = "Naive", GroupLabel2 = "Mixed", 
                      EFDR_M = 0.1, EFDR_D = 0.1, EFDR_R = 0.1)

# Find genes more variable or more expressed in mixed condition
genes.DV <- Test$TableResDisp$GeneName[Test$TableResDisp$ResultDiffResDisp == "Mixed+" |
                                         Test$TableMean$ResultDiffMean == "Mixed+"]
genenames[genes.DV,2]
```

```{r}
# Normalize data
# Offset correction
OffSet <- function(MCMC1, MCMC2){
  OS <- median(rowSums(MCMC1@parameters$mu)/rowSums(MCMC2@parameters$mu))
  OS
}

OS <- OffSet(MCMC.mixed, MCMC.naive)

MCMC.mixed@parameters$mu <- MCMC.mixed@parameters$mu / OS
MCMC.mixed@parameters$phi <- MCMC.mixed@parameters$phi * OS

norm.mixed <- BASiCS_DenoisedCounts(Data.mixed, MCMC.mixed)
norm.naive <- BASiCS_DenoisedCounts(Data.naive, MCMC.naive)
```

In the previous version, we detected Il2 to be DE and DV.

```{r, fig.height=5, fig.width=5}
# Visualize the expression of Il2 - ENSMUSG00000027720
genes <- c("ENSMUSG00000027720")

df <- data.frame(value = c(as.vector(t(norm.mixed[genes,])), as.vector(t(norm.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(norm.mixed)),
                                  rep(genes, each = ncol(norm.naive)))),
                 Condition = factor(c(rep("Mixed", length(as.vector(t(norm.mixed[genes,])))),
                                      rep("Naive", length(as.vector(t(norm.naive[genes,]))))), 
                                    levels = c("Mixed", "Naive"))
                 )
final.jitter <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition), size = 2) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + 
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
final.jitter

ggsave("../../New_Figures_Results/Figure_S5/Fig_S5C_Il2_cata.pdf", final.jitter, width = 7, height = 5)
```

Now, Il2 is boderline significant. 
Alternatively, we can visualize Nfkbid

```{r, fig.height=5, fig.width=5}
# Alternatively, we can use Nfkbid - ENSMUSG00000036931
genes <- c("ENSMUSG00000036931")

df <- data.frame(value = c(as.vector(t(norm.mixed[genes,])), as.vector(t(norm.naive[genes,]))),
                 Genes = factor(c(rep(genes, each = ncol(norm.mixed)),
                                  rep(genes, each = ncol(norm.naive)))),
                 Condition = factor(c(rep("Mixed", length(as.vector(t(norm.mixed[genes,])))),
                                      rep("Naive", length(as.vector(t(norm.naive[genes,]))))), 
                                    levels = c("Mixed", "Naive"))
                 )
final.jitter <- ggplot(data=df, aes(x=Genes, y=log10(value + 1))) + 
  geom_boxplot(outlier.shape = NA, position=position_dodge(width = 0.5), width=0.5, aes(fill=Condition), alpha=0.3) + coord_flip() + 
  geom_jitter( position=position_dodge(width = 0.5), aes(colour=Condition), size = 2) + 
  scale_fill_manual(values = c("seagreen4", "violetred4")) +
  scale_colour_manual(values = c("seagreen4", "violetred4")) + 
  ylab("log10(Expr)") + xlab("Genes") + 
  theme(axis.ticks.y = element_blank(), axis.title.y = element_blank(), axis.line.y = element_blank(),
        strip.background = element_blank(), strip.text.y = element_text(angle = 180))
final.jitter

ggsave("../../New_Figures_Results/Figure_S5/Fig_S5C_cata.pdf", final.jitter, width = 7, height = 5)
```

```{r, fig.height=5, fig.width=10}
# Plot heatmap of DV genes
library(pheatmap)
library(viridis)
pdf("../../New_Figures_Results/Figure_S5/Fig_S5D_cata.pdf", width = 7, height = 12)
pheatmap(log10(norm.mixed[genes.DV,] + 1), color = viridis(100),
         border_color = NA, show_colnames = FALSE, show_rownames = FALSE, 
         annotation_col = data.frame(row.names = colnames(norm.mixed),
                                     activation_state = sapply(colnames(norm.mixed),
                                      function(n){unlist(strsplit(n, "\\."))[4]})))
dev.off()
```